---
title: "analysis_stabilityRD"
output: html_document
date: "2024-09-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r cars}
library(ggplot2)
library(tidyverse)
library(gridExtra)
library(rEDM)
library("parallel")
library(pracma)
library(purrr)
library(signal)
library(ggfortify)
library(data.table)
library(patchwork)
library(codyn)
library(ggrepel)
```

# load, combine and curate dataset
```{r pressure, echo=FALSE}
new_selection_mean <- read_csv("new_selection_mean_simple_div.csv")
load("~/Documents/GitHub/Multifarious-Response-diversity---experiments/Analysis/Data analysis/dens_biomass_mono.RData")
dens_mono<-dens
load("~/Documents/GitHub/Multifarious-Response-diversity---experiments/Analysis/Data analysis/dens_biomass_poly.RData")

#create a day variable from the date variable
earliest_date<-min(dens$date)
days_since_earliest<-as.numeric(dens$date-earliest_date)+1
dens<-dens%>%mutate(day=days_since_earliest,richness=as.factor(richness))

#create a day variable from the date variable for mono_culture
date_format = "%Y-%m-%d"

dens_mono<-dens_mono%>%mutate(date=as.Date(date))

tcl<-dplyr::filter(dens_mono,species_initial%in%c("T","C","L"))
tcl<-mutate(tcl,sample_group="tcl")
earliest_date <- min(tcl$date)
days_since_earliest <- as.numeric(tcl$date - earliest_date) + 1
tcl<-tcl%>%mutate(day=days_since_earliest)

ds<-dplyr::filter(dens_mono,species_initial%in%c("D","S"))
ds<-mutate(ds,sample_group="ds")
earliest_date <- min(ds$date)
days_since_earliest <- as.numeric(ds$date - earliest_date) + 1
ds<-ds%>%mutate(day=days_since_earliest)

p<-dplyr::filter(dens_mono,species_initial%in%c("P"))
p<-mutate(p,sample_group="p")
earliest_date <- min(p$date)
days_since_earliest <- as.numeric(p$date - earliest_date) + 1
p<-p%>%mutate(day=days_since_earliest)

dens_mono<-rbind(tcl,ds,p)


# complete time series 
##  unfortunate steps since sample_ID variable in loaded dens data frame is wrong
#dens<-dens%>%mutate(sample_ID=paste(composition,temperature,nutrients,microcosm,sep="_"))
fill_dd<-expand.grid(sample_ID=unique(dens$sample_ID),day=unique(dens$day))

# get complete ts (26 days * 243 microcosms)
dens_complete<-full_join(fill_dd,dens,join_by(sample_ID,day))
dens_complete<-dens_complete%>%separate(sample_ID, into = c("composition", "temperature", "nutrients", "microcosm"), sep = "_")
dens_complete$biomass[is.na(dens_complete$biomass)]<-0
dens_complete$mean.dens.ml[is.na(dens_complete$mean.dens.ml)]<-0




# get average density and biomass; and also CV
dens_aggr<-dens_complete%>%group_by(composition,nutrients,temperature,microcosm)%>%summarise(
  mean_dens=mean(mean.dens.ml),
  mean_biom=mean(biomass),
  CV=sd(biomass)/mean(biomass),
  CV_dens=sd(mean.dens.ml)/mean(mean.dens.ml)
)

dens_aggr_mono<-dens_mono%>%group_by(species_initial,nutrients,temperature,microcosm)%>%summarise(
  mean_dens=mean(mean.dens.ml),
  mean_biom=mean(biomass),
  CV=sd(biomass)/mean(biomass)
)


# 
dens_aggr<-dens_aggr%>%mutate(nutrients=as.character(nutrients),
                              temperature=paste0(temperature," °C"),
                              nutrients=paste0(nutrients," g/L"))
dens_aggr<-full_join(new_selection_mean,dens_aggr,by=join_by(composition,nutrients,temperature))
```
# Exploration
## First explorative plots 
```{r}
plot_time_series<-ggplot(data=dens,aes(x=day,y=biomass,colour = richness))+
  geom_point()+
  geom_smooth()+
  theme_bw()+
  facet_grid(temperature~nutrients,scales="free_y")


plot_RD_CV<-ggplot(data=dens_aggr,aes(x=mean_divergence,y=log10(1/CV),color=as.factor(richness)))+
  geom_point()+
  geom_smooth(method="lm")

plot_RD_CV_treatment<-ggplot(data=dens_aggr,aes(x=mean_divergence,y=log10(1/CV)))+
  geom_point()+
  geom_smooth(method="lm")+
  facet_grid(temperature~nutrients)

## effects of temperature, nutrients and richness

plot_T<-ggplot(data=dens_aggr,aes(y=log10(1/CV),x=temperature))+
  geom_boxplot()+
  geom_jitter()
plot_N<-ggplot(data=dens_aggr,aes(y=log10(1/CV),x=nutrients))+
  geom_boxplot()+
  geom_jitter()
plot_R<-ggplot(data=dens_aggr,aes(y=log10(1/CV),x=as.factor(richness)))+
  geom_boxplot()+
  geom_jitter()

## look at monoculture

plot_mono_ts<-ggplot(data=dens_mono,aes(x=day,y=biomass,color=species_initial))+
  geom_point()+
  geom_smooth(se=F)+
  facet_grid(temperature~nutrients)


plot_T_mono<-ggplot(data=dens_aggr_mono,aes(y=1/CV,x=as.factor(temperature)))+
  geom_boxplot()+
  geom_jitter()
plot_N_mono<-ggplot(data=dens_aggr_mono,aes(y=1/CV,x=as.factor(nutrients)))+
  geom_boxplot()+
  geom_jitter()

# compare
grid.arrange(plot_T_mono,plot_T,plot_N_mono,plot_N, ncol=2 )
grid.arrange(plot_N_mono,plot_N, ncol=2)

#look at dissimilarity
plot_dissim_CV<-ggplot(data=dens_aggr,aes(x=mean_dissimilarity,y=log10(1/CV),color=as.factor(richness)))+
  geom_point()+
  geom_smooth(method="lm")

plot_DV_dissim<-ggplot(data=dens_aggr,aes(x=mean_dissimilarity,y=mean_divergence))+
  geom_point()

plot_DV_alt_CV<-ggplot(data=dens_aggr,aes(x=mean_divergence_alt,y=log10(1/CV),color=as.factor(richness)))+
  geom_point()+
  geom_smooth(method="lm")

#look at density 


plot_dens_biom<-ggplot(data=dens_aggr,aes(x=mean_biom,y=mean_dens))+
  geom_point()



```

## Identification plots
### classification of noise
```{r}
dd_all_pred<-read.csv("morph_dd_pred.csv")
dd_all_pred_nonoise<-read.csv("morph_dd_pred_nonoise.csv")

names(dd_all_pred_nonoise)[names(dd_all_pred_nonoise)=="predict_spec"]<-"pred_nonoise"
dd_all_pred<-full_join(dd_all_pred,select(dd_all_pred_nonoise,id_new,pred_nonoise),join_by(id_new))

hist_classification <- ggplot(dd_all_pred) +
  geom_histogram(aes(x = pred_nonoise, fill = "Without noise"), stat = "count") +
  geom_histogram(aes(x = predict_spec, fill = "With noise"), stat = "count") +
  facet_wrap(~composition, scale = "free") +
  scale_fill_manual(name = "with/without noise", values = c("Without noise" = "black", "With noise" = "red3"))


dd_percent_noise<-dd_all_pred%>%mutate(noise = predict_spec=="noise")
dd_percent_noise<-dd_percent_noise%>%group_by(composition,nutrients,temperature)%>%summarise(
                                                                       percent_noise= sum(noise)/length(noise))

dd_percent_noise<-dd_percent_noise%>%mutate(nutrients=as.character(nutrients),
                              temperature=paste0(temperature," °C"),
                              nutrients=paste0(nutrients," g/L"))

hist_noise<-ggplot(data=dd_percent_noise)+
  geom_histogram(aes(x=percent_noise))

dist_noise<-ggplot(data=dd_percent_noise)+
  geom_boxplot(aes(y=percent_noise,x=temperature))+
  geom_jitter(aes(y=percent_noise,x=temperature))+
  facet_wrap(~nutrients)

## more noise classified in low nutrient environment -> different morphology and/or lower total numbers and thus more power to stat outliers
## maybe unbalanced training set of classifier -> check later  by using smaller training set


```

### add biomass to identification data frame 
```{r}
load("~/Documents/GitHub/Multifarious-Response-diversity---experiments/Analysis/Data analysis/ciliate_traits.Rdata")


# needs to have id_new variable
ciliate_traits <- ciliate_traits %>%
  mutate(
    # Remove dots from the date
    cleaned_date = gsub("\\.", "", date),
    # Extract the part of id after the underscore
    id_suffix = sub(".*_(.*)", "\\1", id),
    # Combine cleaned_date, id_suffix, and species_initial into a new variable
    id_new = paste0(cleaned_date, id_suffix, composition)
  ) %>%
  # Optionally, remove the intermediate columns to clean up
  select(-cleaned_date, -id_suffix,-new_id)

uniqueN(ciliate_traits$id_new)==nrow(ciliate_traits) # all unique  ;)



id_dd<-full_join(dd_all_pred,select(ciliate_traits,id_new,biomass),join_by("id_new"))


## add day variable

#create a day variable from the date variable

id_dd<-mutate(id_dd,date=as.Date(date,format = "%d.%m.%y"))

earliest_date<-min(id_dd$date)
days_since_earliest<-as.numeric(id_dd$date-earliest_date)+1
id_dd<-id_dd%>%mutate(day=days_since_earliest)

#create a summaised df on microcosm level with each species seperate
# Make sure, that we have n_frames and not N_frames
names(id_dd)[names(id_dd) == "N_frames"] <- "n_frames"

#extrapolation_factor <- 9.301902  # for 16 x magnification 
extrapolation_factor <- 9.828125  # for 25 x magnification 
video_biomass_species <- c( "C", "P", "S","D","L","T")

biomasses <- id_dd %>%
  group_by( day,temperature,nutrients,sample_ID,composition,pred_nonoise) %>% # group  by xxx
  summarize(
    biomass = sum(biomass * n_frames, na.rm = TRUE) / (1 * 125) # if not 3 videos corrections is done below with dens_factor
  ) %>%
  mutate(
    biomass = biomass * extrapolation_factor,
    )



dd_ts_id<-biomasses


plot_ts<-ggplot(dd_ts_id,aes(x=day,y=biomass,color=pred_nonoise))+
  geom_point()+
  geom_smooth(se=F)+
  facet_grid(temperature~nutrients,scale="free")

plot_ts_035_18_21<-ggplot(dplyr::filter(dd_ts_id,pred_nonoise!="noise",temperature=="18-21",nutrients==0.35),aes(x=day,y=biomass,color=pred_nonoise))+
  geom_point()+
  geom_smooth(se=F)+
  facet_wrap(~sample_ID,scale="free")

plot_ts_001_18_21<-ggplot(dplyr::filter(dd_ts_id,pred_nonoise!="noise",temperature=="18-21",nutrients==0.01),aes(x=day,y=biomass,color=pred_nonoise))+
  geom_point()+
  geom_smooth(se=F)+
  facet_wrap(~sample_ID,scale="free")

plot_ts_075_18_21<-ggplot(dplyr::filter(dd_ts_id,pred_nonoise!="noise",temperature=="18-21",nutrients==0.75),aes(x=day,y=biomass,color=pred_nonoise))+
  geom_point()+
  geom_smooth(se=F)+
  facet_wrap(~sample_ID,scale="free")

plot_ts_035_25_28<-ggplot(dplyr::filter(dd_ts_id,pred_nonoise!="noise",temperature=="25-28",nutrients==0.35),aes(x=day,y=biomass,color=pred_nonoise))+
  geom_point()+
  geom_smooth(se=F)+
  facet_wrap(~sample_ID,scale="free")



```

### exploration with incomplete ts

```{r}
#### effects of temperature; nutrient and richness, now with complete time series
incomplete_aggr<-dd_ts_id%>%group_by(composition,nutrients,temperature,sample_ID)%>%summarise(
  mean_biom=mean(biomass),
  CV=sd(biomass)/mean(biomass)
)
incomplete_aggr<-incomplete_aggr%>%mutate(richness=nchar(composition),
                                          nutrients=paste0(nutrients," g/L"),
                                          temperature=paste0(temperature," °C"))

incomplete_aggr<-full_join(new_selection_mean,incomplete_aggr,by=join_by(composition,nutrients,temperature,richness))
incomplete_aggr<-incomplete_aggr%>%mutate(richness=as.factor(richness))

plot_T_complete<-ggplot(data=incomplete_aggr)+
  geom_boxplot(aes(y=log10(1/CV),x=temperature))+
  geom_jitter(aes(y=log10(1/CV),x=temperature))

plot_N_complete<-ggplot(data=incomplete_aggr)+
  geom_boxplot(aes(y=log10(1/CV),x=nutrients))+
  geom_jitter(aes(y=log10(1/CV),x=nutrients,color=richness))

plot_R_complete<-ggplot(data=incomplete_aggr)+
  geom_boxplot(aes(y=log10(1/CV),x=richness))+
  geom_jitter(aes(y=log10(1/CV),x=richness))

plot_CV_divergence_incomp<-ggplot(incomplete_aggr,aes(x=mean_divergence,y=log10(1/CV),color=richness))+
  geom_point()+
  geom_smooth(method="lm")

```



### fill missing time points
```{r}

#fill up missing dates with biomass<-0

fill_dd<-expand.grid(sample_ID=unique(dd_ts_id$sample_ID),day=unique(dd_ts_id$day),pred_nonoise=unique(dd_ts_id$pred_nonoise))
complete_ts<-full_join(fill_dd,dd_ts_id,join_by(sample_ID,day,pred_nonoise))

complete_ts$biomass[is.na(complete_ts$biomass)]<-0
complete_ts<-complete_ts%>%mutate(composition=sub("_.*", "", sample_ID))
complete_ts<-complete_ts %>%
  mutate(temperature = sapply(strsplit(as.character(sample_ID), "_"), function(x) paste(x[3], x[4], sep = "-")))
complete_ts<- mutate(complete_ts,nutrients = gsub(".*Nut(.*?)_.*", "\\1", sample_ID))


# Now remove wrong combinations of composition and predict_spec / pred_nonoise

complete_ts<- complete_ts %>%
  rowwise() %>%
  dplyr::filter(pred_nonoise %in% unlist(strsplit(composition, ""))) %>%
  ungroup()  
complete_ts<-mutate(complete_ts,temperature=as.character(temperature),
                    nutrients=as.character(nutrients),
                    richness=nchar(composition))

plot_complete_ts<-ggplot(complete_ts,aes(x=day,y=biomass,color=pred_nonoise))+
  geom_point()+
  geom_smooth(se=F)+
  facet_wrap(~composition,scale="free")

plot_complete_035_25_28<-ggplot(data=dplyr::filter(complete_ts, temperature=="25-28",nutrients=="0.35"),aes(x=day,y=biomass,color=pred_nonoise))+
  geom_point()+
  geom_smooth(se=F)+
  facet_wrap(~sample_ID,scale="free")
complete_ts<-mutate(complete_ts,richness=nchar(composition))

plot_complete_biomass_richness<-ggplot(data=complete_ts,aes(y=biomass,x=day,color=as.factor(richness)))+
  geom_smooth(method="loess")+
  facet_grid(temperature~nutrients)



```

## Biomass contribution


```{r}
complete_ts<-complete_ts%>%group_by(sample_ID,composition,day)%>%mutate(tot_biomass=sum(biomass))
complete_ts<-complete_ts%>%mutate(biom_contribution=biomass/tot_biomass)

complete_ts<-complete_ts%>%mutate(temperature=paste0(temperature," °C"),
                                      nutrients=paste0(nutrients," g/L"))

plot_complete_035_25_28<-ggplot(data=dplyr::filter(complete_ts, temperature=="25-28 °C",nutrients=="0.35 g/L"),aes(x=day,y=biom_contribution,fill=pred_nonoise))+
  geom_area()+
  scale_fill_brewer(palette = "Set3")+
  facet_wrap(~sample_ID,scale="free")

plot_complete_035_18_21<-ggplot(data=dplyr::filter(complete_ts, temperature=="18-21 °C",nutrients=="0.35 g/L"),aes(x=day,y=biom_contribution,fill=pred_nonoise))+
  geom_area(se=F)+
  scale_fill_brewer(palette = "Set3")+
  facet_wrap(~sample_ID,scale="free")

plot_complete_001_18_21<-ggplot(data=dplyr::filter(complete_ts, temperature=="18-21 °C",nutrients=="0.01 g/L"),aes(x=day,y=biom_contribution,fill=pred_nonoise))+
  geom_area(se=F)+
  scale_fill_brewer(palette = "Set3")+
  facet_wrap(~sample_ID,scale="free")


df_slopes <- read_csv("df_slopes.csv")
names(df_slopes)[names(df_slopes)=="species_initial"]<-"pred_nonoise"

slope_ts<-full_join(select(df_slopes,nutrients,pred_nonoise,temperature,slope),complete_ts)
slope_ts<-slope_ts%>%mutate(w_slope=biom_contribution*slope)

slope_ts<-slope_ts%>%group_by(sample_ID,temperature,nutrients,richness,composition,day,tot_biomass)%>%summarize(sum_w_slopes=abs(sum(w_slope)))

slope_ts<-slope_ts%>%mutate(richness=as.factor(richness))

#divide by 0 total biomass -> NA ... what to do with that?
#slope_ts$sum_w_slopes[is.na(slope_ts$sum_w_slopes)]<-0




##create new variable where it checks, where the last observation =0 is; with complete_ts
aggr_ts <- slope_ts %>%
  group_by( sample_ID) %>%
  arrange(day) %>%
  mutate(
    # Create a flag for non-zero tot_biomass
    non_zero_biomass = tot_biomass != 0,
    # Find the last non-zero day
    last_non_zero_day = ifelse(any(non_zero_biomass), max(day[non_zero_biomass], na.rm = TRUE), NA),
    # Find the first zero day after the last non-zero day
    first_zero_day = ifelse(
      !is.na(last_non_zero_day),
      min(day[!non_zero_biomass & day > last_non_zero_day], na.rm = TRUE),
      NA
    ),
    # Flag for days after the first zero day
    is_after_first_zero_day = ifelse(!is.na(first_zero_day), day > first_zero_day, FALSE)
  ) %>%
  ungroup()

aggr_ts<-aggr_ts%>%mutate(rep_var=sub("_[^_]+$", "", sample_ID))

plot_balance_035_18_21<-ggplot(data=aggr_ts%>%dplyr::filter(nutrients=="0.35 g/L",temperature=="18-21 °C",is_after_first_zero_day==FALSE),aes(x=day,y=sum_w_slopes,group=sample_ID))+
  geom_point()+
  geom_smooth(se=F)+
  facet_wrap(~composition,scale="free")

plot_complete_035_18_21+ plot_balance_035_18_21


ls_ts_microcosm<-aggr_ts%>%mutate(rep_var=sub("_[^_]+$", "", sample_ID))%>%dplyr::filter(is_after_first_zero_day==FALSE,first_zero_day>11)

######
ls_ts_microcosm<-split(ls_ts_microcosm,ls_ts_microcosm$sample_ID)

options(dplyr.summarise.inform = FALSE)

balance_ts<-lapply(ls_ts_microcosm,function(m){
 # Number of days in a window
window_length <- 8
days=unique(m$day)
  

# Initialize an empty list to store the windows
windows_list <- list()

# Ensure that the loop doesn't go out of bounds
for (i in 1:(length(days) - window_length + 1)) {
  # Create a window of length 6
  if ((i + window_length - 1) <= length(days)) {
    window <- days[i:(i + window_length - 1)]
    windows_list[[i]] <- window
  }
}
  
  ts<-lapply(windows_list,function(w){
    df<-m%>%dplyr::filter(day%in%w)
    df<-df%>%group_by(composition,nutrients,temperature,sample_ID)%>%summarise(
  mean_biom=mean(tot_biomass),
  CV=sd(tot_biomass)/mean(tot_biomass),
  sd=sd(tot_biomass),
  avg_w_sumslopes=mean(sum_w_slopes,na.rm = T),
  window_start=min(day),window_end=max(day)
)
    
    return(df)
  })

  ts<-do.call("rbind",ts)
  return(ts)
})

df_balance_ts<-do.call("rbind",balance_ts)
df_balance_ts<-df_balance_ts%>%mutate(richness=nchar(composition)%>%as.factor())


plot_CV_ts<-ggplot(data=df_balance_ts%>%dplyr::filter(nutrients=="0.35 g/L",temperature=="18-21 °C"),aes(x=window_start,y=log10(1/CV)))+
  geom_line()+
  #geom_line(aes(x=window_start,y=avg_w_sumslopes),color="red")+
  facet_wrap(~sample_ID,scale="free_y")

plot_avg_w_sumslopes_ts<-ggplot(data=df_balance_ts%>%dplyr::filter(nutrients=="0.35 g/L",temperature=="18-21 °C"))+
  #geom_line()+
  geom_line(aes(x=window_start,y=log10(avg_w_sumslopes)),color="red")+
  facet_wrap(~sample_ID,scale="free_y")

plot_avg_w_sumslopes_ts+plot_CV_ts

##### Cor within microcosm

plot_cor_sumslope_micro<-ggplot(data=df_balance_ts%>%dplyr::filter(nutrients=="0.35 g/L",temperature=="18-21 °C"),aes(x=log10(avg_w_sumslopes),y=1/CV))+
  geom_point()+
  facet_wrap(~sample_ID)

plot_cor_sumslope<-ggplot(data=df_balance_ts,aes(x=log10(avg_w_sumslopes),y=1/CV,color=richness))+
  geom_point()+
  geom_smooth(method="lm")


df_w_slope_micro<-df_balance_ts%>%dplyr::filter(!is.na(CV))%>%group_by(richness,sample_ID,temperature,nutrients,composition)%>%summarize(cor=cor(avg_w_sumslopes,1/CV))

plot_hist_cor_sumlope_micro<-ggplot(data=df_w_slope_micro)+
  geom_histogram(aes(x=cor),bins=20)

plot_box_cor_sumlope_micro<-ggplot(data=df_w_slope_micro)+
  geom_boxplot(aes(y=cor,x=richness))+
  geom_jitter(aes(y=cor,x=richness))

# cor within days among microcosms

df_w_slope_days_pearson<-df_balance_ts%>%dplyr::filter(!is.na(CV))%>%group_by(window_start)%>%summarize(
  cor=cor.test(x=log10(avg_w_sumslopes),y=1/CV,method="pearson")$estimate,
  uppr=cor.test(x=log10(avg_w_sumslopes),y=1/CV,method="pearson")$conf.int[2],
  lwr=cor.test(x=log10(avg_w_sumslopes),y=1/CV,method="pearson")$conf.int[1],
  pvalue=cor.test(x=log10(avg_w_sumslopes),y=1/CV,method="pearson")$p.value)

plot_ts_cor_pearson<-ggplot(data=df_w_slope_days_pearson,aes(x=window_start,y=cor))+
  geom_line()+
  geom_ribbon(aes(ymin=lwr,ymax=uppr),alpha=0.1)+
  geom_hline(yintercept=0,linetype="dotted")

library(RVAideMemoire)

df_w_slope_days_spearman<-df_balance_ts%>%dplyr::filter(!is.na(CV))%>%group_by(window_start)%>%summarize(
  cor=spearman.ci(avg_w_sumslopes%>%log10(),1/CV)$estimate,
  lwr=spearman.ci(avg_w_sumslopes%>%log10(),1/CV)$conf.int[1],
  uppr=spearman.ci(avg_w_sumslopes%>%log10(),1/CV)$conf.int[2])

plot_ts_cor_spearman<-ggplot(data=df_w_slope_days_spearman,aes(x=window_start,y=cor))+
  geom_line()+
  geom_ribbon(aes(ymin=lwr,ymax=uppr),alpha=0.1)+
  geom_hline(yintercept=0,linetype="dotted")

df_w_slope_days_kendall<-df_balance_ts%>%dplyr::filter(!is.na(CV))%>%group_by(window_start)%>%summarize(
  cor=cor.test(x=avg_w_sumslopes%>%log10(),y=1/CV,method="kendall")$estimate,
  p_value=cor.test(x=avg_w_sumslopes%>%log10(),y=1/CV,method="kendall")$p.value)

plot_ts_cor_kendall<-ggplot(data=df_w_slope_days_kendall,aes(x=window_start,y=cor))+
  geom_line()+
  geom_hline(yintercept=0,linetype="dotted")

# do the same but with richness as a level

df_w_slope_pearson_richness<-df_balance_ts%>%dplyr::filter(!is.na(CV))%>%group_by(window_start,richness)%>%summarize(
  cor=cor.test(x=avg_w_sumslopes%>%log10(),y=1/CV)$estimate,
  uppr=cor.test(x=avg_w_sumslopes%>%log10(),y=1/CV)$conf.int[2],
  lwr=cor.test(x=avg_w_sumslopes%>%log10(),y=1/CV)$conf.int[1])

plot_ts_cor_R<-ggplot(data=df_w_slope_pearson_richness,aes(x=window_start,y=cor))+
  geom_line()+
  geom_ribbon(aes(ymin=lwr,ymax=uppr),alpha=0.1)+
  geom_hline(yintercept=0,linetype="dotted")+
  facet_wrap(~richness)

df_w_slope_days_spearman_R<-df_balance_ts%>%dplyr::filter(!is.na(CV))%>%group_by(window_start,richness)%>%summarize(
  cor=spearman.ci(avg_w_sumslopes%>%log10(),1/CV)$estimate,
  lwr=spearman.ci(avg_w_sumslopes%>%log10(),1/CV)$conf.int[1],
  uppr=spearman.ci(avg_w_sumslopes%>%log10(),1/CV)$conf.int[2])

plot_ts_spear_R<-ggplot(data=df_w_slope_days_spearman_R,aes(x=window_start,y=cor))+
  geom_line()+
  geom_ribbon(aes(ymin=lwr,ymax=uppr),alpha=0.1)+
  geom_hline(yintercept=0,linetype="dotted")+
  facet_wrap(~richness)


##do the same as with correlation but with the slope of an lm

df_w_slope_days_lm<-df_balance_ts%>%dplyr::filter(!is.na(CV))%>%group_by(window_start)%>%summarize(
  slope=summary(lm(1/CV~log10(avg_w_sumslopes)))$coefficients[2,1],
  se=summary(lm(1/CV~log10(avg_w_sumslopes)))$coefficients[2,2])

plot_ts_lm<-ggplot(data=df_w_slope_days_lm,aes(x=window_start,y=slope))+
  geom_line()+
  geom_ribbon(aes(ymin=slope-se,ymax=slope+se),alpha=0.1)+
  geom_hline(yintercept=0,linetype="dotted")
  #facet_wrap(~richness)



#### biomass ts
#take the mean of three replicates 
biomass_ts<-aggr_ts%>%group_by(day,temperature,nutrients,rep_var,richness,composition)%>%summarize(tot_biom=mean(tot_biomass),se_tot_biom=sd(tot_biomass)/sqrt(3))

plot_biomass<-ggplot(data=biomass_ts)+
  #geom_ribbon(aes(x=day,y=mean_tot_biom,fill=richness,color=richness,ymin=mean_tot_biom-se_tot_biom,ymax=mean_tot_biom+se_tot_biom),alpha=0.5)+
  geom_line(method="loess",se=F,aes(x=day,y=tot_biom,group=rep_var,color=richness),size=0.8)+
 
  theme_bw()+
  #geom_text_repel( data=biomass_ts%>%dplyr::filter(day==58,tot_biom!=0),aes(x=day,y=tot_biom,group=composition,label = composition), size = 4,max.overlaps = 20) +
  facet_wrap(temperature~nutrients,scale="free_y")

```

##Exploration completed time series

```{r}

## discard all the rows that are part of time stretches where entire microcosms has (seemingly) gone extinct
complete_aggr<-aggr_ts%>%dplyr::filter(is_after_first_zero_day==FALSE)%>%group_by(composition,nutrients,temperature,sample_ID)%>%summarise(
  mean_biom=mean(tot_biomass),
  CV=sd(tot_biomass)/mean(tot_biomass),
  sd=sd(tot_biomass),
  avg_w_sumslopes=mean(sum_w_slopes,na.rm = T)
)


complete_aggr<-full_join(new_selection_mean,complete_aggr,by=join_by(composition,nutrients,temperature))%>%
  mutate(richness=as.factor(richness))

complete_aggr<-complete_aggr%>%mutate(abs_sum_slope=abs(sum_slopes))


#### effects fidtributions of 1/CV, divergence, dissimilarity

dist_CV_complete<-ggplot(data=complete_aggr,aes(x=log10(1/CV)))+
  geom_histogram(bins=30)

dist_divergence_complete<-ggplot(data=complete_aggr,aes(x=mean_divergence))+
  geom_histogram(bins=15)

dist_alt_divergence_complete<-ggplot(data=complete_aggr,aes(x=mean_divergence_alt))+
  geom_histogram(bins=15)

#too many values with divergence=0

##look at other response diversity

plot_mean_simple_div<-ggplot(data=complete_aggr,aes(x=mean_divergence,y=simple_divergence))+
  geom_point()



dist_dissim_complete<-ggplot(data=complete_aggr,aes(x=mean_dissimilarity))+
  geom_histogram(bins=15)



plot_divergence_dissimilarity<-ggplot(complete_aggr,aes(x=mean_divergence,y=sqrt(mean_dissimilarity)))+
  geom_point()+geom_smooth(method="lm")

cor_dd<-distinct(complete_aggr%>%select(mean_divergence,mean_dissimilarity))
cor_test<-cor.test(cor_dd$mean_divergence, cor_dd$mean_dissimilarity%>%sqrt(), method = "spearman")

plot_divergence_altdiv<-ggplot(complete_aggr,aes(x=mean_divergence,y=mean_divergence_alt))+
  geom_point()+geom_smooth(method="lm")


#### effects of temperature; nutrient and richness, now with complete time series

plot_T_complete<-ggplot(data=complete_aggr)+
  geom_boxplot(aes(y=1/CV,x=temperature))+
  geom_jitter(aes(y=1/CV,x=temperature))

plot_N_complete<-ggplot(data=complete_aggr)+
  geom_boxplot(aes(y=1/CV,x=nutrients))+
  geom_jitter(aes(y=1/CV,x=nutrients,color=richness))

plot_R_complete<-ggplot(data=complete_aggr)+
  geom_boxplot(aes(y=1/CV,x=richness))+
  geom_jitter(aes(y=1/CV,x=richness))

#global CV~divergence

plot_CV_divergence_comp<-ggplot(complete_aggr,aes(x=mean_divergence,y=1/CV, colour = richness))+
  geom_point()+
  geom_smooth(method="lm")

plot_CV_divergence_nut<-ggplot(complete_aggr,aes(x=mean_divergence,y=1/CV,color=nutrients))+
  geom_point()+
  geom_smooth(method="lm")
plot_CV_divergence_temp<-ggplot(complete_aggr,aes(x=mean_divergence,y=log10(1/CV),color=temperature))+
  geom_point()+
  geom_smooth(method="lm")

plot_mean_divergence<-ggplot(complete_aggr,aes(x=mean_divergence_alt,y=1/CV))+
  geom_point()+
  geom_smooth(method="lm")

### simple divergence

plot_CV_simple_divergence<-ggplot(complete_aggr,aes(x=simple_divergence,y=1/CV))+
  geom_point()+
  geom_smooth(method="lm")+
  theme_classic()+
  xlab("recalculated divergence")+
  theme(axis.title.y=element_text(size=16),
        axis.title.x=element_text(size=16))
plot_mean_divergence<-ggplot(complete_aggr,aes(x=mean_divergence,y=1/CV))+
  geom_point()+
  geom_smooth(method="lm")+
  theme_classic()+
  xlab("mean divergence")+
  theme(axis.title.y=element_text(size=16),
        axis.title.x=element_text(size=16))

plot_div_comparison<-plot_mean_divergence+plot_CV_simple_divergence

ggsave("divergence.pdf", plot = plot_div_comparison, width = 10, height = 6)

plot_cor_simple_mean_div<-ggplot(complete_aggr,aes(x=mean_divergence,y=simple_divergence,color=richness))+
  geom_point()+ geom_abline(slope = 1, intercept = 0, color = "black") 




####### response symmetry

plot_CV_abs_sum_slope<-ggplot(complete_aggr,aes(x=log10(abs_sum_slope),y=1/CV,color=richness))+
  geom_point()+
  geom_smooth(method="lm")+theme_bw()

plot_CV_w_sum_slope<-ggplot(complete_aggr,aes(x=log10(avg_w_sumslopes),y=1/CV,color=richness))+
  geom_point()+
  geom_smooth(method="lm")+theme_bw()

plot_cor_abs_sum_div<-ggplot(complete_aggr,aes(x=mean_divergence,y=log10(abs_sum_slope),color=richness))+
  geom_point()
###


#see smaller slope with higher richness -> divergence does only consider the two most "extreme" species in terms of their performance change -> interaction between richness and divergence

#global CV~dissimilatiy

plot_CV_dissim_comp<-ggplot(complete_aggr,aes(x=mean_dissimilarity,y=1/CV))+
  geom_point()+
  geom_smooth(method="lm")



plot_CV_split_comp<-ggplot(complete_aggr,aes(x=mean_divergence,y=1/CV))+
  geom_point(aes(color=richness))+
  geom_smooth(method="lm")+
  facet_grid(temperature~nutrients,scale="free")




plot_CV_split_comp_dissim<-ggplot(complete_aggr,aes(x=mean_dissimilarity%>%sqrt(),y=log10(1/CV)))+
  geom_point(aes(color=richness))+
  geom_smooth(method="lm")+
  facet_grid(temperature~nutrients,scale="free")





```
##Exploration with density

```{r}
load("~/Desktop/Til/community_exp_new/dens_biomass_poly.RData")

#create a day variable from the date variable
earliest_date<-min(dens$date)
days_since_earliest<-as.numeric(dens$date-earliest_date)+1
dens<-dens%>%mutate(day=days_since_earliest,richness=as.factor(richness))
dens<-dens%>%dplyr::filter(!is.na(sample_ID))
#create a day variable from the date variable for mono_culture
date_format = "%Y-%m-%d"

##  unfortunate steps since sample_ID variable in loaded dens data frame is wrong
#dens<-dens%>%mutate(sample_ID=paste(composition,temperature,nutrients,microcosm,sep="_"))
fill_dd<-expand.grid(sample_ID=unique(dens$sample_ID),day=unique(dens$day))
fill_dd<-fill_dd%>%dplyr::filter(!is.na(sample_ID))
# get complete ts (26 days * 243 microcosms)
dens_complete<-dplyr::full_join(fill_dd,dens,join_by(sample_ID,day))

dens_complete$biomass[is.na(dens_complete$biomass)]<-0
dens_complete$mean.dens.ml[is.na(dens_complete$mean.dens.ml)]<-0

# fill out some NAs
dens_complete<-dens_complete%>%mutate(composition=sub("_.*", "", sample_ID))
dens_complete<-dens_complete %>%
  mutate(temperature = sapply(strsplit(as.character(sample_ID), "_"), function(x) paste(x[3], x[4], sep = "-")))
dens_complete<- mutate(dens_complete,nutrients = gsub(".*Nut(.*?)_.*", "\\1", sample_ID))

plot_dens_ts<-ggplot(data=dens_complete,aes(x=day,y=mean.dens.ml,color=composition))+
  geom_point()+
  facet_grid(temperature~nutrients)

dens_aggr<-dens_complete%>%group_by(composition,nutrients,temperature,sample_ID)%>%summarise(
  mean_biom=mean(biomass),
  CV=sd(biomass)/mean(biomass),
  mean_dens=mean(mean.dens.ml),
  CV_dens=sd(mean.dens.ml)/mean(mean.dens.ml)
)


dens_aggr<-dens_aggr%>%mutate(temperature=paste0(temperature," °C"),
                                      nutrients=paste0(nutrients," g/L"))

dens_aggr<-full_join(new_selection_mean,dens_aggr,by=join_by(composition,nutrients,temperature))%>%
  mutate(richness=as.factor(richness))

dens_CV_diver<-ggplot(data=dens_aggr,aes(x=mean_divergence,y=log10(1/CV_dens),color=richness))+
  geom_point()+
  geom_smooth(method="lm")

dens_CV_diver<-ggplot(data=dens_aggr,aes(x=mean_dissimilarity,y=log10(1/CV_dens),color=richness))+
  geom_point()+
  geom_smooth(method="lm")

## without complete time series

dens_aggr_incomplete<-dens%>%group_by(composition,nutrients,temperature,sample_ID)%>%summarise(
  mean_biom=mean(biomass),
  CV=sd(biomass)/mean(biomass),
  mean_dens=mean(mean.dens.ml),
  CV_dens=sd(mean.dens.ml)/mean(mean.dens.ml)
)

dens_aggr_incomplete<-dens_aggr_incomplete%>%mutate(temperature=paste0(temperature," °C"),
                                      nutrients=paste0(nutrients," g/L"))

dens_aggr_incomplete<-full_join(new_selection_mean,dens_aggr_incomplete,by=join_by(composition,nutrients,temperature))%>%
  mutate(richness=as.factor(richness))

dens_CV_diver_incomp<-ggplot(data=dens_aggr_incomplete%>%dplyr::filter(nutrients!="0.01 g/L"),aes(x=mean_divergence,y=log10(1/CV),color=richness))+
  geom_point()+
  geom_smooth(method="lm")

dens_CV_diver_incomp<-ggplot(data=dens_aggr_incomplete,aes(x=mean_dissimilarity,y=log10(1/CV),color=richness))+
  geom_point()+
  geom_smooth(method="lm")


```



# Species interactions
## functions by Uriah
```{r}

# change libSize accordingly (number of time points - max(E) )

best_E_fct <- function(data, target, interactor){
  max1 <- c()
  max2 <- c()
  for(i in 2:5){
    ccm_out <- CCM(dataFrame = data, target = target, column = interactor,libSizes = c(26-i,26-i,1), Tp = -1, E = i, sample = 100)
    # max1[i-1] <- ccm_out[1,2]
    max2[i-1] <- ccm_out[1,3]
  }
  df <- data.frame(target=target,
                   interactor=interactor,
                   Best_E = which.max(max2)+1)
  return(df)
}

CCM_convergence_test <- function(data, target, interactor, libsize, step=3, E, lmin=NULL, lmax=NULL, maxlag=4){
  seeds <- sample(10000:99999, 100, replace = F)
  if(is.null(lmin)){
    lmin <- floor(nrow(data)*0.2)
  }
  lmin <- max(lmin,E+2)
  
  if(is.null(lmax)){
    #lmax <- floor(nrow(data)*0.8) - (maxlag+E-2) 
    lmax <- nrow(data) - (maxlag+E-2)
  }
  
  rhos_bool <- sapply(seeds, function(seed){
    ccm_out <- CCM(dataFrame = data, target = target, column = interactor,
                   libSizes = paste(lmin, lmax, lmax-lmin), Tp = 0, E = E,
                   sample = 1, seed = seed)
    rho_lmin <- ccm_out[1,3]
    rho_lmax <- ccm_out[2,3]
    return(rho_lmin >= rho_lmax)
  })
  
  pvalue <- sum(rhos_bool)/length(rhos_bool)
}

CCM_significance <- function(data, target, interactor, libsize, step=3, E, surrogateMethod="seasonal", T_period=8, rho){
  ts_int <- unlist(data[,interactor])
  ts_tar <- unlist(data[,target])
  if(interactor!="norm_sampleT") {
    surr_interactor = SurrogateData(ts_int, method = surrogateMethod, T_period = T_period,
                                    num_surr = 1000, alpha = 3)
  } else {
    surr_interactor = SurrogateData(ts_int, method = "random_shuffle",
                                    num_surr = 1000, alpha = 3)      
  }
  
  rho_surr <- data.frame(interactor_rho = numeric(1000))
  
  interactor_data = as.data.frame(cbind(data[,1], ts_tar, surr_interactor))
  names(interactor_data) = c("time", target, paste("T", as.character(seq(1, 1000)),	sep = ""))
  
  sapply(1:1000,function(i){
    targetCol = paste("T", i, sep = "") 
    ccm_out = CCM(dataFrame = interactor_data, E = E, Tp = 0, columns = target,
                  target = targetCol, libSizes = paste(libsize, libsize, step), sample = 1)
    col = paste(target, ":", targetCol, sep = "")
    rho_surr$interactor_rho[i] = ccm_out[1, col]
  })
  
  significance <- 1 - ecdf(rho_surr$interactor_rho)(rho)
  return(significance)
}
```

## SMap
```{r}




#create temperature time series patterns
pattern <- c(18, 18, 18, 19.5, 21, 21, 21, 19.5)
T1<-rep(pattern, length.out = 60)
T1<-data.frame(sample_T=T1[unique(complete_ts$day)+1],day=unique(complete_ts$day))

pattern <- c(22, 22, 22, 23.5, 25, 25, 25, 23.5)
T2<-rep(pattern, length.out = 60)
T2<-data.frame(sample_T=T2[unique(complete_ts$day)+1],day=unique(complete_ts$day))

pattern <- c(25, 25, 25, 26.5, 28, 28, 28, 26.5)
T3<-rep(pattern, length.out = 60)
T3<-data.frame(sample_T=T3[unique(complete_ts$day)+1],day=unique(complete_ts$day))


## add temperature time series and perform interpolation and noramlization
## note that total time series goes from sampling day 1 (second day of the experiment) until the 59th day (here 58)-> 1:58


df_S<-split(async_df,async_df$sample_ID)
df_S<-mclapply(df_S,function(df){
  if(unique(df$temperature)=="18-21"){
    dfT<-T1
  }
  else if(unique(df$temperature)=="22-25"){
    dfT<-T2
  }
  else{dfT<-T3}
  
   interpolated <- lapply(unique(df$predict_spec),function(s){
     new_time <- seq(min(df$day), max(df$day), length.out = 26)
     int_sampleT<-interp1(dplyr::filter(df,predict_spec==s)$day, dfT$sample_T, new_time, method = "cubic")
     df_transform<-data.frame(int_biomass=interp1(dplyr::filter(df,predict_spec==s)$day, dplyr::filter(df,predict_spec==s)$biomass, new_time, method = "cubic"),
                original_biomass=dplyr::filter(df,predict_spec==s)$biomass,int_sampleT=int_sampleT,sample_T=dfT$sample_T,day=new_time,predict_spec=rep(s,length(new_time)))
     
    df_transform<-df_transform%>%mutate(trans_biomass=original_biomass^1/4)
    df_transform<-df_transform%>%mutate(resid_biomass=as.vector(lm(data=df_transform,trans_biomass~new_time)$residuals))
    df_transform<-df_transform%>%mutate(norm_trans_biomass=(resid_biomass - mean(resid_biomass)) / sd(resid_biomass),
                                        norm_sampleT=(int_sampleT - mean(int_sampleT)) / sd(int_sampleT),rep_var=dplyr::filter(df,predict_spec==s)$rep_var,
                                        sample_ID=dplyr::filter(df,predict_spec==s)$sample_ID)
    
   })
   df<-do.call("rbind",interpolated)
   df$norm_trans_biomass[is.na(df$norm_trans_biomass)]<-0

   
   return(df)
})

df_S<-do.call("rbind",df_S)

df_S<-split(df_S,df_S$rep_var)

# to test us this
test_df_S<-df_S[23]

#### two questions: inlude T_sample as an interactor? Transorm/interpolate/normalize time series of biomass (and Temperature?)

pairwise_ccm_df_list_filtered<- mclapply(df_S, function(df){
  var_pairs = expand.grid(target=unique(df$pred_nonoise),interactor = append(unique(df$pred_nonoise),"norm_sampleT"))
  species<-unique(df$pred_nonoise)
  
dftemp<-lapply(species,function(s){
  dfs<-df%>%dplyr::filter(pred_nonoise==s)%>%select(day,norm_trans_biomass,norm_sampleT)
  names(dfs)[names(dfs)=="norm_trans_biomass"]<-s
  return(dfs)
})

dftemp<-reduce(dftemp,full_join,by=c("day","norm_sampleT"))

  ############## best embedding dimension based on simplex 

	Best_E_df <- apply(var_pairs,1, function(c){
		best_E_fct(dftemp,c[1],c[2])
	})
	Best_E_df <- do.call("rbind",Best_E_df)
	Best_E_df$ID <- unique(df$ID)
	
  ############## produce prediction skill for every possible interactor-target pair at largest library size using best E from above
	rhos <- apply(Best_E_df, 1, function(r){
		ccm_out <- CCM(dataFrame = dftemp, target = r[1], column = r[2],
									 libSizes = c(2+as.numeric(r[3]),26-as.numeric(r[3]),1), Tp = 0, E = as.numeric(r[3]), sample = 100)
		rho <- ccm_out[26-2-(2*as.numeric(r[3])),3] ###### change to rho from largest library size (example: if E=5:  21-5)
		if(rho<0) return(data.frame(rho=rho,comment="negative"))
		else {return(data.frame(rho=rho,comment="positive"))}
	})

	Best_E_df <- cbind(Best_E_df,do.call("rbind",rhos))

  ############## get significance of prediction skill using surrogate time series
    significances <- apply(Best_E_df,1,function(r){
      pvalue<-CCM_significance(data=dftemp,target=r[1],interactor=r[2],libsize =26-as.numeric(r[3]) ,surrogateMethod="seasonal",T_period=8,E=as.numeric(r[3]),rho=r[4])
})
  ############## get significance for convergence criteria 
    convergence<- apply(Best_E_df,1,function(r){
  pvalue<-CCM_convergence_test(data=dftemp, target=r[1],interactor=r[2], libsize=26-as.numeric(r[3]), step=3, E=as.numeric(r[3]), lmin=NULL, lmax=NULL, maxlag=4)
  })
  
	pairwise_ccm_df <- cbind(Best_E_df,significances,convergence)

	
significant_pairwise_ccm_df <- pairwise_ccm_df %>%
	dplyr::filter(significances <= 0.05, rho >0, convergence<=0.05)

significant_pairwise_ccm_df<-split(significant_pairwise_ccm_df,significant_pairwise_ccm_df$target)

list_map<-lapply(significant_pairwise_ccm_df,function(x){

  interactor<-x$interactor
  lib=c(1,26)

  SMap(dataFrame=dftemp,target=unique(x$target),columns=interactor,E=length(interactor),theta=5,lib=lib,pred=lib,embedded=T)
})

list_map

}, mc.cores = detectCores()-2)

#save(pairwise_ccm_df_list, file="CCM_interpolated.RData")
#save(pairwise_ccm_df_list_n, file="CCM_rawdata.RData")
#save(pairwise_ccm_df_list_trans, file="CCM_transformed.RData")
#save(pairwise_ccm_df_list_trans_noTemp, file="CCM_transformed_noTemp.RData")
save(pairwise_ccm_df_list_filtered, file="CCM_filtered.RData")

```


## plot Interactions


```{r}

load("~/Documents/GitHub/Multifarious-Response-diversity---experiments/Analysis/Data analysis/CCM_filtered.RData")
load("~/Documents/GitHub/Multifarious-Response-diversity---experiments/Analysis/Data analysis/CCM_transformed.RData")

# unpack ccm list
results_list <- list()

# Loop through each Sample_ID in the list
for(sample_id in names(pairwise_ccm_df_list_trans)) {
  
  # Loop through each target in the Sample_ID element
  for(target in names(pairwise_ccm_df_list_trans[[sample_id]])) {
    
    # Extract the coefficients data frame
    coeff_df <- pairwise_ccm_df_list_trans[[sample_id]][[target]]$coefficients
    
    # Remove the C0 column
    coeff_df <- coeff_df[ , !names(coeff_df) %in% "C0"]
    
    # Get the additional coefficient column names (after day)
    extra_cols <- setdiff(names(coeff_df), "day")
    
    # Reshape the data: gather all extra columns into one column
    reshaped_df <- tidyr::gather(coeff_df, key = "coefficient_name", value = "coefficient_value", extra_cols)
    
    # Add columns for Sample_ID and target
    reshaped_df$Sample_ID <- sample_id
    reshaped_df$target <- target
    
    # Add the reshaped data to the results list
    results_list[[paste0(sample_id, "_", target)]] <- reshaped_df
  }
}

# Combine all the results into one data frame
interactions_df <- do.call(rbind, results_list)

#create temperature and nutrients variable
interactions_df<-interactions_df %>%
  mutate(temperature = sapply(strsplit(as.character(Sample_ID), "_"), function(x) paste(x[3], x[4], sep = "-")))
interactions_df<- mutate(interactions_df,nutrients = gsub(".*Nut(.*?)_.*", "\\1", Sample_ID))
interactions_df<-interactions_df%>%mutate(composition=sub("_.*", "", Sample_ID))
#interactions_df<-interactions_df%>%mutate(richness=as.factor(nchar(composition)),coefficient_value=abs(coefficient_value))

all_partial_deriv<-ggplot(data=interactions_df,aes(x=log10(coefficient_value)))+
  geom_histogram()+
  facet_wrap(nutrients~temperature,scale="free_y")
  #geom_jitter(aes(cex=20))


#exclude intraspecific interaction
interactions_df<-dplyr::filter(interactions_df,coefficient_name!="∂C/∂C" & coefficient_name!="∂D/∂D" & coefficient_name!="∂L/∂L" &
                          coefficient_name!="∂S/∂S" & coefficient_name!="∂P/∂P")

# exclude interaction of temperature
interactions_df<-dplyr::filter(interactions_df,coefficient_name!="∂C/∂norm_sampleT" & coefficient_name!="∂D/∂norm_sampleT" & coefficient_name!="∂L/∂norm_sampleT" &
                          coefficient_name!="∂S/∂norm_sampleT" & coefficient_name!="∂P/∂norm_sampleT")




plot_int_ts<-ggplot(interactions_df%>%dplyr::filter(nutrients==0.75),aes(x=day,y=coefficient_value,color=coefficient_name))+
  geom_line()+
  facet_wrap(~Sample_ID,scale="free")


mean_int_species<-interactions_df%>%group_by(Sample_ID,coefficient_name,temperature,nutrients,composition,target)%>%summarize(mean_interaction=mean(abs(coefficient_value),na.rm = T))
mean_int_species<-mutate(mean_int_species,richness=as.factor(nchar(composition)))

mean_int_species<-mean_int_species%>%group_by(Sample_ID,temperature,nutrients,composition,target,richness)%>%summarize(number_interactions=length(coefficient_name))

all_partial_deriv<-ggplot(data=mean_int_species,aes(y=number_interactions,x=richness))+
  geom_boxplot()+
  geom_jitter(height=0.1,aes(color=nutrients))+
  facet_wrap(~target,scale="free")
  #geom_jitter(aes(cex=20))

all_partial_richness<-ggplot(data=mean_int_species,aes(y=log10(mean_interaction),x=richness))+
  geom_boxplot()+
  facet_grid(nutrients~temperature,scale="free_x")+
  geom_jitter()

# mean_interaction_df<- interactions_df%>%group_by(Sample_ID,day,temperature,nutrients,composition)%>%reframe(mean_interaction=sum(coefficient_value)/((nchar(composition)^2) - nchar(composition)))

mean_interaction_df<- interactions_df%>%group_by(Sample_ID,day,temperature,nutrients,composition)%>%reframe(mean_interaction=mean(coefficient_value),sum_interaction=sum(abs(coefficient_value)),mean_interaction_noabs=mean(coefficient_value))


plot_mean_interactions<-ggplot(data=mean_interaction_df,aes(x=day,y=mean_interaction,color=composition))+
  geom_line()+
  facet_grid(temperature~nutrients,scale="free")

overall_mean_interaction<-mean_interaction_df%>%group_by(Sample_ID,temperature,nutrients,composition)%>%reframe(mean_sum_interactions=mean(sum_interaction,na.rm=T),mean_interaction=mean(mean_interaction,na.rm=T))



## join interactions with main df (complete_aggr)
names(overall_mean_interaction)[names(overall_mean_interaction)=="Sample_ID"]<-"sample_ID"

int_aggr<-inner_join(dplyr::select(overall_mean_interaction,mean_interaction,sample_ID,mean_sum_interactions),complete_aggr,join_by(sample_ID))

## distributions

plot_dist_interaction<-ggplot(data=int_aggr,aes(x=mean_interaction))+
  geom_histogram(bins=30)
plot_dist_interaction_log<-ggplot(data=int_aggr,aes(x=log10(mean_interaction_strength)))+
  geom_histogram(bins=30)


plot_dist_CV<-ggplot(data=int_aggr,aes(x=1/CV))+
  geom_histogram(bins=30)
plot_dist_CV_log<-ggplot(data=int_aggr,aes(x=log10(1/CV)))+
  geom_histogram(bins=30)
grid.arrange(plot_dist_CV,plot_dist_CV_log)


### should we log transform 1/CV ?

plot_CV_int_mean<-ggplot(data=int_aggr,aes(x=log10(mean_interaction_strength),y=1/CV,color=richness))+
  geom_point()+
  geom_smooth(method="lm")
  #facet_grid(temperature~nutrients,scale="free")

plot_CV_int_sum<-ggplot(data=int_aggr,aes(x=log10(mean_sum_interactions),y=1/CV,color=richness))+
  geom_point()+
  geom_smooth(method="lm")
  #facet_grid(temperature~nutrients,scale="free")



plot_divergence_int<-ggplot(data=int_aggr,aes(y=log10(mean_interaction_strength),x=mean_divergence))+
  geom_point()+
  geom_smooth(method="lm")
# ->no relationship between interaction and divergence

plot_dissimilarity_int<-ggplot(data=int_aggr,aes(y=log10(mean_interaction_strength),x=mean_dissimilarity%>%sqrt(),color=richness))+
  geom_point()+
  geom_smooth(method="lm")

plot_richness_int<-ggplot(data=int_aggr,aes(y=log10(mean_interaction_strength),x=richness))+
  geom_boxplot()+
  geom_jitter()


plot_temp_int<-ggplot(data=int_aggr,aes(y=log10(mean_interaction_strength),x=temperature))+
  geom_boxplot()+
  geom_jitter(aes(color=nutrients))

plot_nut_int<-ggplot(data=int_aggr,aes(y=log10(mean_interaction_strength),x=nutrients))+
  geom_boxplot()+
  geom_jitter()


##### look at CV~divergence with all alternative values

plot_divergence_R_CV<-ggplot(int_aggr,aes(y=log10(1/CV),x=mean_divergence,color=richness))+
  geom_point()+
  geom_smooth(method="lm")

plot_int_simple_div_CV<-ggplot(int_aggr,aes(y=log10(1/CV),x=simple_divergence))+
  geom_point()+
  geom_smooth(method="lm")


plot_abs_sum_slopes_CV<-ggplot(int_aggr,aes(y=log10(1/CV),x=log10(abs_sum_slope)))+
  geom_point()+
  geom_smooth(method="lm")

#interactions vs r_div

plot_abs_sum_slopes_int<-ggplot(int_aggr,aes(y=log10(mean_interaction_strength),x=log10(abs_sum_slope)))+
  geom_point()+
  geom_smooth(method="lm")

plot_simple_divergence_int<-ggplot(int_aggr,aes(y=log10(mean_interaction),x=simple_divergence))+
  geom_point()+
  geom_smooth(method="lm")






plot_dissim_R_CV<-ggplot(int_aggr,aes(y=log10(1/CV),x=mean_dissimilarity,color=richness))+
  geom_point()+
  geom_smooth(method="lm")



####### what about all the microcosms that show no interactions ##############

no_int_aggr<-complete_aggr%>%dplyr::filter(!sample_ID %in% int_aggr$sample_ID)

plot_divergence_no_int<-ggplot(no_int_aggr,aes(y=log10(1/CV),x=mean_divergence,color=richness))+
  geom_point()+
  geom_smooth(method="lm")

# even more pronounced interaction between richness and mean divergence for species where we have not found evidence for interactions...
# does this mean that species interactions is not the root cause of that interactive effect of richness?

plot_richness_no_int<-ggplot(data=no_int_aggr,aes(y=log10(1/CV),x=richness))+
  geom_boxplot()+
  geom_jitter()

# richness seems to have a negative effect on stability

plot_temp_no_int<-ggplot(data=no_int_aggr,aes(y=log10(1/CV),x=temperature))+
  geom_boxplot()+
  geom_jitter()

# so does T
plot_nut_no_int<-ggplot(data=no_int_aggr,aes(y=log10(1/CV),x=nutrients))+
  geom_boxplot()+
  geom_jitter()

# and nutrients seem to have a positive effect

# set all communities with 0 interactions detected to = = (or rather +1)
 no_int_aggr <- no_int_aggr%>%mutate(mean_interaction_strength=0,
                                     mean_sum_interactions=0)
 
 total_int_aggr<-full_join(no_int_aggr,int_aggr)
 
 total_int_aggr<-total_int_aggr%>%mutate(mean_interaction_strength=mean_interaction_strength+1,
                                     mean_sum_interactions=mean_sum_interactions+1)
 
 
 
 plot_CV_int_mean_tot<-ggplot(data=total_int_aggr,aes(x=log10(mean_interaction_strength),y=log10(1/CV),color=richness))+
  geom_point()+
  geom_smooth(method="lm")
  #facet_grid(temperature~nutrients,scale="free")
 
 plot_CV_int_sum_tot<-ggplot(data=total_int_aggr,aes(x=log10(mean_sum_interactions),y=log10(1/CV),color=richness))+
  geom_point()+
  geom_smooth(method="lm")
  #facet_grid(temperature~nutrients,scale="free")
 
 #-> richness 3 and 4 no longer negative slopes...

```



##Extinctions

```{r}

complete_ts_micro_level<-complete_ts%>%group_by(sample_ID,day,temperature,nutrients,composition)%>%summarise(sum_biomass=sum(biomass),mean_biomass=mean(biomass))
complete_ts_micro_level<-complete_ts_micro_level%>%mutate(richness=nchar(composition)%>%as.factor())






plot_biomass_incomplete<-ggplot(data=dd_ts_id)+
  geom_line(method="loess",se=F,aes(x=day,y=biomass,group=sample_ID,color=richness),colour="black",size=0.8,alpha=0.2)+
  geom_smooth(aes(x=day,y=biomass,color=pred_nonoise),se=F,method="loess")+
  theme_bw()+
  facet_wrap(temperature~nutrients,scale="free_y")


plot_biomass_complete<-ggplot(data=complete_ts)+
  geom_line(method="loess",se=F,aes(x=day,y=biomass,group=sample_ID),colour="black",size=0.8,alpha=0.2)+
  geom_smooth(aes(x=day,y=biomass,color=pred_nonoise),se=F,method="loess")+
  theme_bw()+
  facet_wrap(temperature~nutrients,scale="free_y")

##create new variable where it check, where the last observation =0 is; with complete_ts
complete_ts <- complete_ts %>%
  group_by(pred_nonoise, sample_ID) %>%
  arrange(day) %>%
  mutate(
    # Create a flag for non-zero biomass
    non_zero_biomass = biomass != 0,
    # Find the last non-zero day
    last_non_zero_day = ifelse(any(non_zero_biomass), max(day[non_zero_biomass], na.rm = TRUE), NA),
    # Find the first zero day after the last non-zero day
    first_zero_day = ifelse(
      !is.na(last_non_zero_day),
      min(day[!non_zero_biomass & day > last_non_zero_day], na.rm = TRUE),
      NA
    ),
    # Flag for days after the first zero day
    is_after_first_zero_day = ifelse(!is.na(first_zero_day), day >= first_zero_day, FALSE)
  ) %>%
  ungroup()
complete_ts<-complete_ts %>%
  group_by(sample_ID, pred_nonoise) %>%
  mutate(relative_zero_biomass = sum(biomass == 0)/richness) %>%
  ungroup()

extinction_df<-complete_ts%>%group_by(sample_ID,nutrients,temperature,composition,richness,pred_nonoise)%>%summarize(last_obs=unique(last_non_zero_day),tot_zero=unique(relative_zero_biomass))

extinction_df$last_obs[is.na(extinction_df$last_obs)]<-0

plot_last_observation<-ggplot(data=extinction_df,aes(x=last_obs,y=as.factor(temperature)))+
  geom_boxplot()+
  geom_jitter(aes(color=as.factor(richness)))+
  facet_wrap(~pred_nonoise)+
  xlab("last observation [day]")

hist_extinctions<-ggplot(data=extinction_df,aes(y=last_obs,x=as.factor(richness)))+
  geom_boxplot()+
  geom_jitter()
  

## total extinction day on level of microcosm
extinction_df_aggr<-extinction_df%>%group_by(sample_ID,nutrients,temperature,composition,richness)%>%summarize(extinction_microcosm=max(last_obs),tot_zeroes=sum(tot_zero),na.rm = T)%>%ungroup()

plot_extinction<-ggplot(data=extinction_df_aggr,aes(y=extinction_microcosm,x=richness%>%as.factor()))+
  geom_boxplot()+
  geom_jitter()+
  facet_grid(temperature~nutrients)+
  ylab("last day before extinction")

extinction_df_aggr<-select(extinction_df_aggr,sample_ID,extinction_microcosm,tot_zeroes)

extinction_df_aggr<-full_join(extinction_df_aggr%>%select(sample_ID,extinction_microcosm,tot_zeroes),complete_aggr)

plot_ex_div<-ggplot(data=extinction_df_aggr,aes(x=mean_divergence,y=extinction_microcosm))+
  geom_point()+
  geom_smooth(method="lm")

plot_ex_CV_complete<-ggplot(data=extinction_df_aggr,aes(y=log10(1/CV),x=extinction_microcosm))+
  geom_point()+
  geom_smooth(method="lm")+
  xlab("day of last observation")

plot_last_obs_richness<-ggplot(data=extinction_df_aggr,aes(x=richness,y=extinction_microcosm))+
  geom_boxplot()+
  geom_jitter()


plot_tot_zeroes_richness<-ggplot(data=extinction_df_aggr,aes(x=richness,y=tot_zeroes))+
  geom_boxplot()+
  geom_jitter()

plot_tot_zeroes_nut_richness<-ggplot(data=extinction_df_aggr,aes(x=nutrients,y=tot_zeroes,color=richness))+
  geom_boxplot()+
  geom_jitter()
# low nutrients have most zeroes -> the least stable


plot_tot_zeroes_CV<-ggplot(data=extinction_df_aggr,aes(y=log10(1/CV),x=tot_zeroes))+
  geom_point()+
  geom_smooth(method="lm")+
  xlab("tot number of zeroes")

# the more zero observations, the less stable 
  

# add CV from incomplete ts

extinction_df_aggr<-full_join(extinction_df_aggr,dens_aggr_incomplete%>%select(sample_ID,CV),join_by(sample_ID))
names(extinction_df_aggr)[extinction_df_aggr%>%names=="CV.x"]<-"CV_complete"
names(extinction_df_aggr)[extinction_df_aggr%>%names=="CV.y"]<-"CV_incomplete"

plot_ex_CV_incomplete<-ggplot(data=extinction_df_aggr,aes(y=log10(1/CV_incomplete),x=extinction_microcosm,color=richness))+
  geom_point()+
  geom_smooth(method="lm")+
  xlab("day of last observation")

#now it is the opposite...

#look at standard deviation and mean
plt_sd_mean<-ggplot(data=extinction_df_aggr,aes(x=sd,y=mean))+
  geom_point()

#CV seams reasonable-> higher mean associated with higher sd

##only look at microcosms that did not go extinct

plot_58_CV_div<-ggplot(data=extinction_df_aggr%>%dplyr::filter(extinction_microcosm==58),aes(x=mean_divergence,y=log10(1/CV_complete),color=richness))+
  geom_point()+
  geom_smooth(method="lm")


# look at extinctions with interactions

extinction_df_int<-inner_join(int_aggr,extinction_df_aggr)

plot_int_ext<-ggplot(data=extinction_df_int,aes(y=extinction_microcosm,x=log10(interaction_strength)))+
  geom_point()
  
plot_int_tot_zeroes<-ggplot(data=extinction_df_int,aes(y=tot_zeroes,x=log10(interaction_strength),color=richness))+
  geom_point()+
  geom_smooth(method="lm")





```


## Asnychrony

```{r}

async_df<-complete_ts%>%mutate(rep_var=sub("_[^_]+$", "", sample_ID))


async_Gross <- async_df %>% group_by(composition,sample_ID) %>%  
  do(synchrony = synchrony(., "day", "pred_nonoise", 
                              "biomass", metric = "Gross", replicate.var = "rep_var"))
async_Gross<-mutate(async_Gross,synchrony=synchrony$synchrony)

async_aggr<-full_join(async_Gross,complete_aggr,join_by(sample_ID,composition))

plot_asynch_CV<-ggplot(data=async_aggr,aes(x=synchrony,y=1/CV,color=richness))+
  geom_point()+
  geom_smooth(method="lm")+theme_bw()

plot_asynch_mean_div<-ggplot(data=async_aggr,aes(y=synchrony,x=log10(abs_sum_slope),color=richness))+
  geom_point()+
  geom_smooth(method="lm")

#-> higher divergence leads to asynchrony in richness=2, resulting in a positive relationship between divergence and tmp stability for richness 2. The opposite happens for the higher richness levels

#look at other divergences

plot_asynch_simple_div<-ggplot(data=async_aggr,aes(y=synchrony,x=simple_divergence,color=richness))+
  geom_point()+
  geom_smooth(method="lm")

plot_asynch_sum_slopes<-ggplot(data=async_aggr,aes(y=synchrony,x=log10(abs_sum_slope),color=richness))+
  geom_point()+
  geom_smooth(method="lm")




### interactions

async_int<-full_join(async_Gross,int_aggr,join_by(sample_ID,composition))

plot_asynch_int<-ggplot(data=async_int,aes(y=synchrony,x=log10(mean_interaction_strength),color=richness))+
  geom_point()+
  geom_smooth(method="lm")


#-> asynchrony goes down with interaction strength (slightly more pronounced with Loreau asynchrony)

plot_asynch_int_b<-ggplot(data=async_int,aes(y=synchrony,x=log10(mean_sum_interactions)))+
  geom_point()+
  geom_smooth(method="lm")

#-> opposite for sum of interactions? (no trend for Loreau method)
```


## Interactions combined microcosms

```{r}
test_df_S<-df_S[23]

rep_CCM<-mclapply(test_df_S,function(df){
  var_pairs = expand.grid(target=unique(df$predict_spec),interactor = append(unique(df$predict_spec),"norm_sampleT"))
  species<-unique(df$predict_spec)
  
dftemp<-lapply(species,function(s){
  dfs<-df%>%dplyr::filter(predict_spec==s)%>%dplyr::select(day,norm_trans_biomass,norm_sampleT,sample_ID)
  names(dfs)[names(dfs)=="norm_trans_biomass"]<-s
  dfs<-split(dfs,dfs$sample_ID)
  dfs<-lapply(dfs,function(df){
    dna<-rbind(df,rep(NA,ncol(df)))
  return(dna)
  })
  dfs<-do.call("rbind",dfs)
  dfs<-dfs%>%mutate(position=1:nrow(dfs))
  dfs<-dfs%>%dplyr::select(-c(sample_ID))
  return(dfs)
})

dftemp<-reduce(dftemp,full_join,by=c("position","day","norm_sampleT"))
dftemp<-dftemp[1:(nrow(dftemp)-1),]

#### find best E
maxE_test<-round(sqrt(26)) 

best_E_df <- lapply(species,function(t){
  A<-dftemp%>%pull(t[1])
  
  max1 <- c()
  max2 <- c()
  for(i in 2:maxE_test){
    ccm_out <- SSR_pred_boot(A=A,B=A,E=i)
  # max1[i-1] <- ccm_out[1,2]
    max2[i-1] <- ccm_out$rho
  }
    
  if(length(max2)==0||all(is.na(max2))){
    return(data.frame(tar=t[1],Best_E=NA,rho=NA))
  }else{return(data.frame(tar=t[1],
                   Best_E = which.max(max2)+1,
                   rho=max(max2)))}
})
best_E_df<-do.call("rbind",best_E_df)
best_E_df<-best_E_df%>%dplyr::filter(!is.na(Best_E))

var_pairs<-var_pairs%>%dplyr::filter(target%in%best_E_df$tar,
                                     interactor%in%best_E_df$tar)

### Pairwise CCM

ccm_all<-apply(var_pairs,1,function(v){
  best_E<-dplyr::filter(best_E_df,tar==v["target"])$Best_E
  A<-dftemp%>%pull(v["interactor"])
  B<-dftemp%>%pull(v["target"])
  
  ccm_out<-CCM_boot(A=A,B=B,E=best_E)
  
  ccm_out$target<-v["target"]
  ccm_out$interactor<-v["interactor"]
  rho_df<-data.frame(max_rho=max(ccm_out$rho))

  ccm_out$rho_df<-rho_df
  return(ccm_out)
})

### covergence testing

causl<-apply(var_pairs,1,function(p){
  target=p["target"]
  interactor=p["interactor"]
  
  
  selected_lists <- Filter(Negate(is.null), lapply(ccm_all, function(x) {
  if (x$target == target & x$interactor == interactor ) {
    return(x)
  }}))
    
  p <- 1 - sum(selected_lists[[1]]$FULLinfo[1, ] < selected_lists[[1]]$FULLinfo[nrow(selected_lists[[1]]$FULLinfo), 
    ], na.rm = T)/ncol(selected_lists[[1]]$FULLinfo)
    
  rho<-(0<selected_lists[[1]]$rho_df$max_rho)
  ddf<-data.frame(pvalue=p,rho_larger_0=rho,target=target,interactor=interactor)

### test of surrogates
  
  ## 1 create surrogate data
     # remove NAs
    dfnna<-dftemp%>%dplyr::filter(!is.na(day))
    ts_int <- unlist(dfnna[,interactor])
    
    ts_tar <- unlist(dftemp%>%select(all_of(target)))
    
    # create surrogates of interactor ts
    if(interactor!="norm_sampleT") {
    surr_interactor = SurrogateData(ts_int, method = "seasonal", T_period = 8,
                                    num_surr = 1000, alpha = 3)
  } else {
    surr_interactor = SurrogateData(ts_int, method = "random_shuffle",
                                    num_surr = 1000, alpha = 3)      
  }
    
    # add back NAs 
    
    na_row <- rep(NA, ncol(surr_interactor))

# Split the matrix and insert NA rows
mat_with_na <- rbind(
  surr_interactor[1:26, ],  # First 26 rows
  na_row,       # Insert NA row
  surr_interactor[27:52, ], # Next 26 rows
  na_row,       # Insert NA row
  surr_interactor[53:nrow(surr_interactor), ] # Remaining rows
)
    
    
    interactor_data = as.data.frame(cbind( dftemp[,1],ts_tar, mat_with_na))
    
  names(interactor_data) = c("time", target, paste("T", as.character(seq(1, 1000)),	sep = ""))
  
  rho_surr <- data.frame(interactor_rho = numeric(1000))
  best_E<-dplyr::filter(best_E_df,tar==target)$Best_E

  sapply(1:1000,function(i){
    targetCol = paste("T", i, sep = "")
    
    
    B_ts<-interactor_data[,targetCol]
    A_ts<-interactor_data[,target]
    
   

    ccm_s = CCM_boot(A=A_ts,B=B_ts,E=best_E,iterations = 10)
    
    rho_surr$interactor_rho[i] = max(unlist(ccm_s$rho))
  })
    significance <- 1 - ecdf(rho_surr$interactor_rho)(unlist(selected_lists[[1]]$rho_df$max_rho))
   return(significance)
   #ddf<-cbind(ddf,sur_p=significance)
})  
  return(causl)
  
})
  
#ccm_df<-do.call("rbind",causl)
  
  

  
  
  
  
  
  ### prepare for s_map
  
  
  significant_pairwise_ccm_df <- ccm_df %>%
	dplyr::filter(rho_larger_0 ==T, pvalue<=0.05,sur_p<=0.05)

  significant_pairwise_ccm_df<-split(significant_pairwise_ccm_df,significant_pairwise_ccm_df$target)
  
  
  
  ### smap
  smap<-lapply(significant_pairwise_ccm_df,function(r){
  
  best_E<-dplyr::filter(best_E_df,target==unique(r$target))$Best_E
  interactor<-r$interactor
  lib=c("1 26","28 53", "55 80")
  pred<-c(c("1 26","28 53", "55 80"))
  

   sss<-lapply(pred,function(p){

      s<-SMap(dataFrame=dftemp,E=best_E,target=unique(r$target),columns=r$interactor,theta=5,lib=lib,pred=p,embedded=T)
      return(s)
    })
     return(sss)
   })
        

return(causl)
}, mc.cores = detectCores()-2)

```

# morphology
```{r}
ts_morph<-id_dd%>%group_by(sample_ID,pred_nonoise,predict_spec,day,temperature,nutrients,composition)%>%summarize(mean_net_speed=mean(net_speed),
                                                                                                     mean_mean_area=mean(mean_area))

plot_mean_netspeed<-ggplot(data=ts_morph,aes(x=day,y=mean_net_speed,color=pred_nonoise))+
  geom_point()+geom_smooth()+
  facet_grid(temperature~nutrients)

plot_mean_boxplot<-ggplot(data=ts_morph,aes(x=nchar(composition)%>%as.factor(),y=mean_net_speed))+
  geom_boxplot()+geom_jitter()+
  facet_grid(temperature~nutrients)

plot_mean_netspeed<-ggplot(data=ts_morph,aes(x=day,y=mean_mean_area,color=pred_nonoise))+
  geom_point()+geom_smooth()+
  facet_grid(temperature~nutrients)

plot_mean_boxplot<-ggplot(data=ts_morph,aes(x=nutrients%>%as.factor(),y=mean_mean_area))+
  geom_boxplot()+geom_jitter()+
  facet_wrap(~predict_spec)

```


# Random slopes

```{r}

slopes<-rnorm(1000,mean=0,sd=0.1)

mean_s<-mean(df_slopes$slope)
sd_s<-sd(df_slopes$slope)

r2_balance<-lapply(1:1000,function(c){
  slopes<-rnorm(2,mean=mean_s,sd=sd_s)
  balance<-abs(mean(slopes))%>%log10()
  
  balance<-data.frame(RD=balance,richness=2)
  return(balance)
})
r2_balance<-do.call("rbind",r2_balance)

r3_balance<-lapply(1:1000,function(c){
  slopes<-rnorm(3,mean=mean_s,sd=sd_s)
  balance<-abs(mean(slopes))%>%log10()
  
  balance<-data.frame(RD=balance,richness=3)
  return(balance)
})
r3_balance<-do.call("rbind",r3_balance)

r4_balance<-lapply(1:1000,function(c){
  slopes<-rnorm(4,mean=mean_s,sd=sd_s)
  balance<-abs(mean(slopes))%>%log10()
  
  balance<-data.frame(RD=balance,richness=4)
  return(balance)
})
r4_balance<-do.call("rbind",r4_balance)

r10_balance<-lapply(1:1000,function(c){
  slopes<-rnorm(10,mean=mean_s,sd=sd_s)
  balance<-abs(mean(slopes))%>%log10()
  
  balance<-data.frame(RD=balance,richness=10)
  return(balance)
})
r10_balance<-do.call("rbind",r10_balance)



balance<-rbind(r3_balance,r2_balance,r4_balance,r10_balance)
balance<-balance%>%mutate(richness=as.factor(richness))

dist_balance<-ggplot(data=balance,aes(x=RD,fill=richness))+
  geom_density(aes(alpha=0.1))
  

means<-balance%>%group_by(richness)%>%summarize(mean_RD=mean(RD))


##-> the higher the richness, the less balanced on average a random community ... its the opposite if one takes the abs(mean(slopes))

#-> taking mean / weighted sum of slopes should lead to a diversity effect on temporal stability

exp_means<-data.frame(exp_means=c(mean(dplyr::filter(complete_aggr,richness==2)$avg_w_sumslopes%>%log10()),
mean(dplyr::filter(complete_aggr,richness==3)$avg_w_sumslopes%>%log10()),
mean(dplyr::filter(complete_aggr,richness==4)$avg_w_sumslopes%>%log10())),richness=c(2,3,4)%>%as.factor())

df_means<-full_join(means[1:3,],exp_means,join_by(richness))

```



# interesting plots

```{r}
## biomass over time

plot_biomass
### => richness =2 more biomass

## response diversity vs temporal stability

plot_CV_divergence_comp+plot_asynch_CV+plot_asynch_mean_div


#### effects of temperature; nutrient and richness

plot_T_complete
plot_N_complete
plot_R_complete

### dissimilarity

plot_CV_dissim_comp


### interactions

plot_CV_int_mean+plot_CV_int_sum

plot_divergence_int

plot_richness_int
plot_temp_int
plot_nut_int


### new response div metrics

plot_CV_simple_divergence

plot_CV_abs_sum_slope+plot_CV_w_sum_slope


```

