---
title: "Untitled"
author: "Til Hämmig"
date: "2024-09-27"
output: html_document
---

```{r setup, include=FALSE}
library("ggplot2")
library("tidyverse")
library("gridExtra")
#library(rEDM)
library("parallel")
library("pracma")
library("purrr")
library("signal")
library("ggfortify")
library("data.table")
library("patchwork")
library("codyn")
library("ggrepel")
library("lme4")
library("lmerTest")
library("MuMIn")
library("RColorBrewer")
library("broom")
library("relaimpo")
library("lavaanPlot")
#devtools::install_github("canankarakoc/r_package_EDMhelper/EDMhelper")

```


```{r load packages}
new_selection_mean <- read_csv("new_selection_mean_simple_div.csv")
load("~/Documents/GitHub/Multifarious-Response-diversity---experiments/Analysis/Data analysis/dens_biomass_poly.RData")

dd_all_pred<-read.csv("morph_dd_pred.csv")
dd_all_pred_nonoise<-read.csv("morph_dd_pred_nonoise.csv")

load("~/Documents/GitHub/Multifarious-Response-diversity---experiments/Analysis/Data analysis/ciliate_traits.Rdata")

df_slopes <- read_csv("df_slopes.csv")
```

## Data cleaning / mutating / transforming


```{r Data cleaning, echo=FALSE}
earliest_date<-min(dens$date)
days_since_earliest<-as.numeric(dens$date-earliest_date)+1
dens<-dens%>%mutate(day=days_since_earliest,richness=as.factor(richness))

#create a day variable from the date variable for mono_culture
date_format = "%Y-%m-%d"

# data for hist_identification
names(dd_all_pred_nonoise)[names(dd_all_pred_nonoise)=="predict_spec"]<-"pred_nonoise"
dd_all_pred<-full_join(dd_all_pred,dplyr::select(dd_all_pred_nonoise,id_new,pred_nonoise),join_by(id_new))

hist_classification <- ggplot(dd_all_pred) +
  geom_histogram(aes(x = pred_nonoise, fill = "Without noise"), stat = "count") +
  geom_histogram(aes(x = predict_spec, fill = "With noise"), stat = "count") +
  facet_wrap(~composition, scale = "free") +
  scale_fill_manual(name = "with/without noise", values = c("Without noise" = "black", "With noise" = "red3"))

```


## Histo classification
```{r histo_identification_noise}

hist_classification
```
## Biomass

```{r biomass calcuations,echo=F}
# needs to have id_new variable
ciliate_traits <- ciliate_traits %>%
  mutate(
    # Remove dots from the date
    cleaned_date = gsub("\\.", "", date),
    # Extract the part of id after the underscore
    id_suffix = sub(".*_(.*)", "\\1", id),
    # Combine cleaned_date, id_suffix, and species_initial into a new variable
    id_new = paste0(cleaned_date, id_suffix, composition)
  ) %>%
  # Optionally, remove the intermediate columns to clean up
  dplyr::select(-cleaned_date, -id_suffix,-new_id)

uniqueN(ciliate_traits$id_new)==nrow(ciliate_traits) # all unique  ;)

id_dd<-full_join(dd_all_pred,dplyr::select(ciliate_traits,id_new,biomass),join_by("id_new"))


## add day variable

#create a day variable from the date variable

id_dd<-mutate(id_dd,date=as.Date(date,format = "%d.%m.%y"))

earliest_date<-min(id_dd$date)
days_since_earliest<-as.numeric(id_dd$date-earliest_date)+1
id_dd<-id_dd%>%mutate(day=days_since_earliest)

#create a summaised df on microcosm level with each species seperate
# Make sure, that we have n_frames and not N_frames
names(id_dd)[names(id_dd) == "N_frames"] <- "n_frames"

#extrapolation_factor <- 9.301902  # for 16 x magnification 
extrapolation_factor <- 9.828125  # for 25 x magnification 
video_biomass_species <- c( "C", "P", "S","D","L","T")

biomasses <- id_dd %>%
  group_by( day,temperature,nutrients,sample_ID,composition,predict_spec) %>% # group  by xxx
  summarize(
    biomass = sum(biomass * n_frames, na.rm = TRUE) / (1 * 125) # if not 3 videos corrections is done below with dens_factor
  ) %>%
  mutate(
    biomass = biomass * extrapolation_factor,
    )



dd_ts_id<-biomasses

#fill up missing dates with biomass<-0

fill_dd<-expand.grid(sample_ID=unique(dd_ts_id$sample_ID),day=unique(dd_ts_id$day),predict_spec=unique(dd_ts_id$predict_spec))
complete_ts<-full_join(fill_dd,dd_ts_id,join_by(sample_ID,day,predict_spec))

complete_ts$biomass[is.na(complete_ts$biomass)]<-0
complete_ts<-complete_ts%>%mutate(composition=sub("_.*", "", sample_ID))
complete_ts<-complete_ts %>%
  mutate(temperature = sapply(strsplit(as.character(sample_ID), "_"), function(x) paste(x[3], x[4], sep = "-")))
complete_ts<- mutate(complete_ts,nutrients = gsub(".*Nut(.*?)_.*", "\\1", sample_ID))

# Now remove wrong combinations of composition and predict_spec / predict_spec

complete_ts<- complete_ts %>%
  rowwise() %>%
  dplyr::filter(predict_spec %in% unlist(strsplit(composition, ""))) %>%
  ungroup()  
complete_ts<-mutate(complete_ts,temperature=as.character(temperature),
                    nutrients=as.character(nutrients),
                    richness=nchar(composition))

complete_ts<-complete_ts%>%group_by(sample_ID,composition,day)%>%mutate(tot_biomass=sum(biomass))
complete_ts<-complete_ts%>%mutate(biom_contribution=biomass/tot_biomass)

complete_ts<-complete_ts%>%mutate(temperature=paste0(temperature," °C"),
                                      nutrients=paste0(nutrients," g/L"))


# introduce slopes of 
names(df_slopes)[names(df_slopes)=="species_initial"]<-"predict_spec"

slope_ts<-full_join(dplyr::select(df_slopes,nutrients,predict_spec,temperature,slope),complete_ts)
slope_ts<-slope_ts%>%mutate(w_slope=biom_contribution*slope,
                            sign=sign(slope))

slope_ts<-slope_ts%>%group_by(sample_ID,temperature,nutrients,richness,composition,day,tot_biomass)%>%summarize(
  sum_w_slopes=abs(sum(w_slope)),
                   mean_abs_slope=mean(abs(slope)),
  sum_abs_slope=sum(abs(slope)),
  abs_sum_slope=abs(sum(slope)),
  symmetry=abs(sum(sign)))


slope_ts<-slope_ts%>%mutate(richness=as.factor(richness))


##create new variable where it checks, where the last observation =0 is; with complete_ts
aggr_ts <- slope_ts %>%
  group_by( sample_ID) %>%
  arrange(day) %>%
  mutate(
    # Create a flag for non-zero tot_biomass
    non_zero_biomass = tot_biomass != 0,
    # Find the last non-zero day
    last_non_zero_day = ifelse(any(non_zero_biomass), max(day[non_zero_biomass], na.rm = TRUE), NA),
    # Find the first zero day after the last non-zero day
    first_zero_day = ifelse(
      !is.na(last_non_zero_day),
      min(day[!non_zero_biomass & day > last_non_zero_day], na.rm = TRUE),
      NA
    ),
    # Flag for days after the first zero day
    is_after_first_zero_day = ifelse(!is.na(first_zero_day), day > first_zero_day, FALSE)
  ) %>%
  ungroup()

aggr_ts<-aggr_ts%>%mutate(rep_var=sub("_[^_]+$", "", sample_ID))

biomass_ts<-aggr_ts%>%group_by(day,temperature,nutrients,rep_var,richness,composition)%>%summarize(tot_biom=mean(tot_biomass),se_tot_biom=sd(tot_biomass)/sqrt(3))


```

### tot biomass plot

```{r plot_biomass,echo=F}
plot_biomass<-ggplot(data=biomass_ts)+
  #geom_ribbon(aes(x=day,y=mean_tot_biom,fill=richness,color=richness,ymin=mean_tot_biom-se_tot_biom,ymax=mean_tot_biom+se_tot_biom),alpha=0.5)+
  geom_line(method="loess",se=F,aes(x=day,y=tot_biom,group=rep_var,color=richness),size=0.8)+
 
  theme_bw()+
  #geom_text_repel( data=biomass_ts%>%dplyr::filter(day==58,tot_biom!=0),aes(x=day,y=tot_biom,group=composition,label = composition), size = 4,max.overlaps = 20) +
  facet_wrap(temperature~nutrients,scale="free_y")
plot_biomass
```
## Main Results 
```{r main results}
## discard all the rows that are part of time stretches where entire microcosms has (seemingly) gone extinct
#%>%dplyr::filter(is_after_first_zero_day==FALSE)

complete_aggr<-aggr_ts%>%group_by(composition,nutrients,temperature,sample_ID)%>%reframe(
  avg_w_sumslopes=mean(sum_w_slopes,na.rm = T),
  abs_sum_slope=mean(abs_sum_slope),
  magnitude=mean(mean_abs_slope),
  symmetry=mean(symmetry),
  sum_abs_slope=mean(sum_abs_slope),
  CV=sd(tot_biomass)/mean(tot_biomass))

complete_aggr<-full_join(new_selection_mean,complete_aggr,by=join_by(composition,nutrients,temperature))%>%
  mutate(richness=as.factor(richness))



complete_aggr<-complete_aggr%>%mutate(stability=1/CV)


```

### Effect of T, N and R
```{r boxplots TNR}
#### effects of temperature; nutrient and richness, now with complete time series

plot_T_complete<-ggplot(data=complete_aggr)+
  geom_boxplot(aes(y=log10(abs_sum_slope),x=temperature))+
  geom_jitter(aes(y=log10(abs_sum_slope),x=temperature,color=richness))

plot_N_complete<-ggplot(data=complete_aggr)+
  geom_boxplot(aes(y=1/CV,x=nutrients))+
  geom_jitter(aes(y=1/CV,x=nutrients,color=richness))

plot_R_complete<-ggplot(data=complete_aggr)+
  geom_boxplot(aes(y=1/CV,x=richness))+
  geom_jitter(aes(y=1/CV,x=richness))

plot_R_complete/
(plot_T_complete+
plot_N_complete)

```


###Effect RD
```{r effect RD}

plot_CV_abs_sum_slope<-ggplot(data=complete_aggr,aes(x=log10(abs_sum_slope),y=1/CV,color=nutrients))+
  geom_point()+
  geom_smooth(method="lm",aes(x=log10(abs_sum_slope),y=1/CV,color=nutrients))+theme_bw()+
  geom_smooth(se=F, aes(x=log10(abs_sum_slope),y=1/CV),method = "lm",color="black")+
  facet_wrap(~temperature)

plot_CV_w_sum_slope<-complete_aggr%>%ggplot(aes(x=log10(avg_w_sumslopes),y=1/CV,color=nutrients))+
  geom_point()+
  geom_smooth(method="lm",aes(x=log10(avg_w_sumslopes),y=1/CV,color=nutrients))+theme_bw()+
 geom_smooth(se=F,aes(x=log10(avg_w_sumslopes),y=1/CV),method = "lm",color="black")

plot_CV_abs_sum_slope+plot_CV_w_sum_slope
```

### Cor sum vs. weighted_sum
```{r Cor sum weighted_sum}

plot_cor<-ggplot(data=complete_aggr,aes(x=log10(abs_sum_slope),y=avg_w_sumslopes%>%log10(
)))+
  geom_point(aes(color=richness))+
   geom_abline(slope = 1, intercept = 0, color = "black")

plot_cor

plot_cor_div_balance<-ggplot(data=complete_aggr,aes(x=log10(abs_sum_slope),y=simple_divergence))+
  geom_point(aes(color=richness))

df_no_0_div<-complete_aggr%>%filter(simple_divergence!=0)

cor.test(df_no_0_div$simple_divergence,df_no_0_div$abs_sum_slope)
```


### Mixed effect model

```{r mixed effect model}



model_1 <- lmer(
     formula = log10(1/CV) ~ log10(abs_sum_slope)+richness +temperature+nutrients+(1 | composition),data = complete_aggr)

anova(model_1,type=3)
check_model(model_1)
summary(model_1)
vif(model_1)
options(na.action=na.fail)


### partial out temperature and nutrients

library(nnet)
#temp_resid <- multinom(temperature ~ log10(abs_sum_slope),data=complete_aggr)$residuals

temp_resid<-lm(data=complete_aggr,log10(abs_sum_slope)~temperature)$residuals

nut_resid<-lm(data=complete_aggr,log10(avg_w_sumslopes)~nutrients)$residuals
balance_resid<-lm(data=complete_aggr,log10(avg_w_sumslopes)~temperature+nutrients)$residuals


new_df<-cbind(complete_aggr,temp_resid,nut_resid=nut_resid,balance_resid)

model_1_a <- lmer(
     formula = log10(stability) ~ temp_resid*richness+nutrients+(1 | composition),data = new_df)

library(performance)
anova(model_1_a)
summary(model_1_a)
vif(model_1_a)
check_model(model_1_a)
#selection_model_1<-dredge(model_1)







model_2 <- lmer(
     formula = 1/CV ~ log10(avg_w_sumslopes)+richness + temperature+  nutrients + (1 | composition),data = complete_aggr)


check_model(model_2)
anova(model_2,type=3)
summary(model_2)
selection_model_2<-dredge(model_2)


model_lm<-lm(data=complete_aggr, log10(stability)~log10(abs_sum_slope) + richness+temperature+nutrients)

par(mfrow=c(2,2))
plot(model_lm)
check_model(model_lm)

summary(model_lm)
anova(model_lm)



```

## Linear model


```{r}

lm_full<-lm(data=complete_aggr, log10(stability)~log10(abs_sum_slope)+richness+temperature+nutrients)

lm_full_div<-lm(data=complete_aggr, log10(stability)~simple_divergence+richness+temperature+nutrients)

anova(lm_full,lm_full_div)


summary(lm_full_div)
anova(lm_full)
check_model(lm_full)


# Calculate relative importance
rel_imp <- calc.relimp(lm_full_w, type = "lmg", rela = TRUE)



lm_full_w<-lm(data=complete_aggr, log10(stability)~log10(avg_w_sumslopes)+richness+temperature+nutrients)

summary(lm_full_w)
anova(lm_full_w)
check_model(lm_full_w)

anova(lm_full_w,lm_full)
```




## model comparisons
### model_mixed_001_1821

```{r model_mixed_001_1821}


model_mixed_001_1821 <- lmer(
     formula = stability ~ log10(sum_abs_slope)+richness  + (1 | composition),
     data = complete_aggr%>%filter(nutrients=="0.01 g/L",temperature=="18-21 °C"))

check_model(model_mixed_001_1821)


summary(model_mixed_001_1821)
anova(model_mixed_001_1821)



```

```{r model_mixed_035_1821}


model_mixed_035_1821 <- lmer(
     formula = stability ~ log10(sum_abs_slope)+richness  + (1 | composition),
     data = complete_aggr%>%filter(nutrients=="0.35 g/L",temperature=="18-21 °C"))

check_model(model_mixed_035_1821)


summary(model_mixed_035_1821)
anova(model_mixed_035_1821)



```
```{r model_mixed_075_1821}


model_mixed_075_1821 <- lmer(
     formula = stability ~ log10(sum_abs_slope)+richness  + (1 | composition),
     data = complete_aggr%>%filter(nutrients=="0.75 g/L",temperature=="18-21 °C"))

check_model(model_mixed_075_1821)


summary(model_mixed_075_1821)
anova(model_mixed_075_1821)



```

```{r model_mixed_001_2225}


model_mixed_001_2225 <- lmer(
     formula = stability ~ log10(sum_abs_slope)+richness  + (1 | composition),
     data = complete_aggr%>%filter(nutrients=="0.01 g/L",temperature=="22-25 °C"))

check_model(model_mixed_001_2225)


summary(model_mixed_001_2225)
anova(model_mixed_001_2225)



```

```{r model_mixed_035_2225}


model_mixed_035_2225 <- lmer(
     formula = stability ~ log10(sum_abs_slope)+richness  + (1 | composition),
     data = complete_aggr%>%filter(nutrients=="0.35 g/L",temperature=="22-25 °C"))

check_model(model_mixed_035_2225)


summary(model_mixed_035_2225)
anova(model_mixed_035_2225)



```
```{r model_mixed_075_2225}


model_mixed_075_2225 <- lmer(
     formula = stability ~ log10(sum_abs_slope)+richness  + (1 | composition),
     data = complete_aggr%>%filter(nutrients=="0.75 g/L",temperature=="22-25 °C"))

check_model(model_mixed_075_2225)


summary(model_mixed_075_2225)
anova(model_mixed_075_2225)



```
```{r model_mixed_001_2528}


model_mixed_001_2528 <- lmer(
     formula = stability ~ log10(sum_abs_slope)+richness  + (1 | composition),
     data = complete_aggr%>%filter(nutrients=="0.01 g/L",temperature=="25-28 °C"))

check_model(model_mixed_001_2528)


summary(model_mixed_001_2528)
anova(model_mixed_001_2528)



```
```{r model_mixed_035_2528}


model_mixed_035_2528 <- lmer(
     formula = stability ~ log10(sum_abs_slope)+richness  + (1 | composition),
     data = complete_aggr%>%filter(nutrients=="0.35 g/L",temperature=="25-28 °C"))

check_model(model_mixed_035_2528)


summary(model_mixed_035_2528)
anova(model_mixed_035_2528)



```
### linear models
```{r lm_001_1821}


lm_001_1821<-lm(data=complete_aggr%>%filter(nutrients=="0.01 g/L",temperature=="18-21 °C"),
                log10(stability)~log10(abs_sum_slope)+richness)


check_model(lm_001_1821)
summary(lm_001_1821)
anova(lm_001_1821)
```



```{r lm_035_1821}


lm_035_1821<-lm(data=complete_aggr%>%filter(nutrients=="0.35 g/L",temperature=="18-21 °C"),
                log10(stability)~log10(abs_sum_slope)+richness)


check_model(lm_035_1821)
summary(lm_035_1821)
anova(lm_035_1821)
```
```{r lm_075_1821}


lm_075_1821<-lm(data=complete_aggr%>%filter(nutrients=="0.75 g/L",temperature=="18-21 °C"),
                log10(stability)~log10(abs_sum_slope)+richness)


check_model(lm_075_1821)
summary(lm_075_1821)
anova(lm_075_1821)
```
```{r lm_001_2225}


lm_001_2225<-lm(data=complete_aggr%>%filter(nutrients=="0.01 g/L",temperature=="22-25 °C"),
                log10(stability)~log10(abs_sum_slope)+richness)


check_model(lm_001_2225)
summary(lm_001_2225)
anova(lm_001_2225)
```
```{r lm_035_2225}


lm_035_2225<-lm(data=complete_aggr%>%filter(nutrients=="0.35 g/L",temperature=="22-25 °C"),
                log10(stability)~log10(abs_sum_slope)+richness)


check_model(lm_035_2225)
summary(lm_035_2225)
anova(lm_035_2225)
```

```{r lm_075_2225}


lm_075_2225<-lm(data=complete_aggr%>%filter(nutrients=="0.75 g/L",temperature=="22-25 °C"),
                log10(stability)~log10(abs_sum_slope)+richness)


check_model(lm_075_2225)
summary(lm_075_2225)
anova(lm_075_2225)
```
```{r lm_001_2528}


lm_001_2528<-lm(data=complete_aggr%>%filter(nutrients=="0.01 g/L",temperature=="22-25 °C"),
                log10(stability)~log10(abs_sum_slope)+richness)


check_model(lm_001_2528)
summary(lm_001_2528)
anova(lm_001_2528)
```

###check random effect


```{r model_mix_random_tn}
model_mix_random_tn <- lmer(
     formula = log10(stability) ~ log10(sum_abs_slope)+richness + (1 | temperature)+(1 | nutrients),
     data = complete_aggr)
check_model(model_mix_random_tn)

summary(model_mix_random_tn)
anova(model_mix_random_tn)



```
### check composition 


```{r}
lm_comp_balance<-lm(data=complete_aggr%>%dplyr::filter(nutrients=="0.01 g/L",temperature=="18-21 °C"),log10(abs_sum_slope)~composition)


check_model(lm_comp_balance)
summary(lm_comp_balance)


lm_comp_balance<-lm(data=complete_aggr,log10(abs_sum_slope)~composition+nutrients*temperature)
summary(lm_comp_balance)
```


## Asynchrony

```{r asynchrony calc.,echo=F}

#filter out completely extinct microcosms


async_df<-complete_ts%>%mutate(rep_var=sub("_[^_]+$", "", sample_ID))


async_Gross <- async_df %>% group_by(composition,sample_ID) %>%  
  do(synchrony_Gross = synchrony(., "day", "predict_spec", 
                              "biomass", metric = "Gross"))
async_Gross<-mutate(async_Gross,synchrony_Gross=synchrony_Gross%>%unlist())

async_Loroeau <- async_df %>% group_by(composition,sample_ID) %>%  
  do(synchrony_L = synchrony(., "day", "predict_spec", 
                              "biomass", metric = "Loreau"))
async_Loroeau<-mutate(async_Loroeau,synchrony_Loreau=synchrony_L%>%unlist())

async_aggr<-full_join(async_Gross,complete_aggr,join_by(sample_ID,composition))

async_aggr<-full_join(async_aggr,async_Loroeau,join_by(sample_ID,composition))

async_aggr_f<-async_aggr%>%dplyr::filter(nutrients!="0.01 g/L")
async_aggr<-async_aggr%>%mutate(rep_var=sub("_[^_]+$", "", sample_ID))
```


### Plots
```{r async plots,echo=F}
plot_asynch_CV_G<-ggplot(data=async_aggr,aes(x=log10(synchrony_Gross+1),y=log10(1/CV),color=nutrients))+
  geom_point()+
  geom_smooth(method="lm")+theme_bw()

plot_asynch_CV_L<-ggplot(data=async_aggr,aes(x=synchrony_Loreau,y=1/CV,color=nutrients))+
  geom_point()+
  geom_smooth(method="lm")+theme_bw()




plot_asynch_CV_G+
plot_asynch_CV_L
```


```{r async cor}
plot_asynch_L_RD<-ggplot(data=async_aggr,aes(y=synchrony_Loreau,x=log10(avg_w_sumslopes),color=nutrients))+
  geom_point()+
  geom_smooth(method="lm")+theme_bw()
cor.test(async_aggr$synchrony_Loreau,async_aggr$avg_w_sumslopes)

plot_asynch_G_RD<-ggplot(data=async_aggr,aes(y=synchrony_Gross,x=log10(abs_sum_slope),color=nutrients))+
  geom_point()+
  geom_smooth(method="lm")+theme_bw()

cor.test(async_aggr$synchrony_Gross,async_aggr$avg_w_sumslopes)
cor.test(async_aggr$synchrony_Loreau,async_aggr$avg_w_sumslopes)

plot_asynch_L_RD+plot_asynch_G_RD
# -> no relationship between synchrony and RD...

model_asyn_CV<-lm(data=async_aggr%>%dplyr::filter(!is.na(synchrony_Gross)),log10(stability)~log10(synchrony_Gross+1))


summary(model_asyn_CV)
anova(model_asyn_CV)
```
### Models
#### Asynchrony Gross with interaction

```{r model inc. asynchrony}
lm_async<-lm(data=async_aggr%>%dplyr::filter(!is.na(synchrony_Gross),nutrients!="0.01 g/L"),log10(1+synchrony_Gross)~log10(abs_sum_slope))

check_model(lm_async)
summary(lm_async)
```

#### Comparison to basic model

```{r model_3/2 comparison }
#### model 2 for comparison (without asynchrony)
model_2 <- lmer(
     formula = I(1 / CV) ~ log10(avg_w_sumslopes)*richness + temperature+  nutrients + (1 | composition),data = async_aggr%>%dplyr::filter(!is.na(synchrony_Gross)))

anova(model_2,model_3)
```
####Comparison to simple linear model

```{r model_3/lm comparison}
### simple linear model for comparison
lm_async<-lm(data=async_aggr%>%dplyr::filter(!is.na(synchrony_Gross),nutrients!="0.01 g/L"),
             synchrony_Gross~log10(abs_sum_slope)+richness+nutrients+temperature)

check_model(lm_async)
summary(lm_async)
anova(lm_async)
```

## SEM and balance sim

```{r interaction RD async}

# simulate random slopes


slopes<-rnorm(1000,mean=0,sd=0.1)

mean_s<-mean(df_slopes$slope)
sd_s<-sd(df_slopes$slope)

RD_df<-lapply(c(2,3,4,10),function(r){
  
  RD<-lapply(1:1000,function(c){
  slopes<-rnorm(r,mean=0,sd=sd_s)
  balance<-abs(mean(slopes))
  simple_divergence<-(max(slopes)-min(slopes)-abs(abs(max(slopes))-abs(min(slopes))))/(max(slopes)-min(slopes))
  adj_divergence<-abs(sum(slopes))/sum(abs(slopes))
  avg_magnitude<-mean(abs(slopes))
  RD_df<-data.frame(balance=balance,simple_divergence=simple_divergence,richness=r,adj_divergence=1-adj_divergence,avg_magnitude=avg_magnitude)
  
  return(RD_df)
})
  RD<-do.call("rbind",RD)
})

RD_df<-do.call("rbind",RD_df)


plot<-ggplot(data=RD_df,aes(x=balance,color=as.factor(richness)))+
  geom_density()

plot_sim_div_bal<-ggplot(data=RD_df,aes(x=adj_divergence,y=balance,color=avg_magnitude))+
  geom_point()+
  facet_wrap(~richness)


plot_div_balance<-ggplot(data=comp_aggr,aes(x=1-rel_compensation,y=avg_w_sumslopes,color=mean_abs_slope))+
  geom_point()+
  facet_wrap(~richness)

plot_sim_div_bal/plot_div_balance



# Ensure the data is ungrouped before applying transformations
sem_aggr <- async_aggr %>%
  ungroup() %>%  # Ensure there is no grouping
  mutate(
    log_abs_sum_slope = log10(abs_sum_slope),
    stability = log10(1 / CV),
    richness = as.numeric(richness),
    temperature=temperature,
    log_magnitude=log10(magnitude),
    sqrt_symm=sqrt(symmetry),
    log_sum_abs=log10(sum_abs_slope),
    RD=abs_sum_slope/sum_abs_slope,
    log_syn_Gross=log10(1+synchrony_Gross),
    symmetry_scaled=(symmetry/as.numeric(richness)),
    asyn_Gross= (-synchrony_Gross),
    log_asyn_Gross=log10((-synchrony_Gross)+1),
    log_sum_abs_slope=log10(sum_abs_slope)
    #Keep it as an ordered factor
  )

quantile(sem_aggr$log_abs_sum_slope)


sem_aggr$nutrients[sem_aggr$nutrients=="0.01 g/L"]<-0.01
sem_aggr$nutrients[sem_aggr$nutrients=="0.35 g/L"]<-0.35
sem_aggr$nutrients[sem_aggr$nutrients=="0.75 g/L"]<-0.75

sem_aggr$temperature[sem_aggr$temperature=="18-21 °C"]=1
sem_aggr$temperature[sem_aggr$temperature=="22-25 °C"]=2
sem_aggr$temperature[sem_aggr$temperature=="25-28 °C"]=3


#correct balance for magnitude
# the effect of magnitude goes down, the more divergence there is

# Load the lavaan package
library(lavaan)


model1A <- '
  stability ~ asyn_Gross+temperature+nutrients+symmetry_scaled+log_sum_abs_slope
  
  asyn_Gross ~ nutrients+richness+log_sum_abs_slope+symmetry_scaled


'
modl<-lm(data=sem_aggr,log_abs_sum_slope~symmetry_scaled+log_sum_abs_slope)


model1B <- '
  stability ~ asyn_Gross
  +temperature
  +nutrients
  +log_abs_sum_slope+
  richness
  
  asyn_Gross ~ log_abs_sum_slope
  +nutrients
  +richness
'



model1C <- '
  stability ~ asyn_Gross+temperature+c("c1","c2","c1")*richness+log_abs_sum_slope
  
  asyn_Gross ~ log_abs_sum_slope+c("d1","d2","d3")*richness
  
  c1==0
  d2==0
'




# Define the SEM model using lavaan syntax
 model2 <- '
  # Direct effects
  stability ~ asyn_Gross+log_magnitude
  
  # Indirect effect via synchrony
  asyn_Gross ~ symmetry


 
'
 


# Fit the model
fit <- sem(model2, data = sem_aggr%>%dplyr::filter(!is.na(synchrony_Gross),nutrients!="0.01 g/L"))


fit1A <- sem(model1A, estimator="MLM",meanstructure = TRUE,data = sem_aggr%>%dplyr::filter(!is.na(synchrony_Gross)))

fit1B <- sem(model1B, estimator="MLM",meanstructure = TRUE,data = sem_aggr%>%dplyr::filter(!is.na(synchrony_Gross)))

fit1C <- sem(model1C, estimator="MLM",meanstructure = TRUE,group="nutrients", data = sem_aggr%>%dplyr::filter(,!is.na(synchrony_Gross)))



 ### model B slightly better
AICc(fit1C,fit1B) ### model B and C almost the same

anova(fit1B,fit1C)
# Summarize the results
summary(fit1B, standardized = TRUE, fit.measures = TRUE)

modificationindices(fit1A)


lavaanPlot(model = fit1A, node_options = list(shape = "box", fontname = "Helvetica"), edge_options = list(color = "grey"), coefs = TRUE, sig = .05)

```
## interactions

```{r stat averaging}

interactions_df <- read_csv("interactions_df_w_T.csv")

int_aggr<-full_join(interactions_df,async_aggr,join_by(sample_ID))
int_aggr$mean_int_strength[is.na(int_aggr$mean_int_strength)]=0
int_aggr$mean_interaction[is.na(int_aggr$mean_interaction)]=0
int_aggr$sum_int_strength[is.na(int_aggr$sum_int_strength)]=0

M_int<-lm(data=int_aggr,log10(stability)~log10(abs_sum_slope)+nutrients+temperature+richness+mean_int_strength)
anova(M_int,lm_full)



plot_int_syn<-ggplot(data=int_aggr,aes(y=synchrony_Gross,x=mean_int_strength,color=log10(abs_sum_slope)))+
  geom_point()+
  geom_smooth(method="lm")





### above rep level
rep_aggr<-int_aggr%>%group_by(rep_var.y,richness,temperature,nutrients)%>% dplyr::reframe(log_mean_fun_balance=log10(mean(abs_sum_slope)),
                                                                                     mean_syn_Gross=mean(synchrony_Gross,na.rm=T),
                                                                                     mean_int_rep_strength=mean(mean_int_strength),
                                                                                     log_mean_stab=log10(mean(1/CV)))
                                                                              

M_int<-lm(data=rep_aggr,log_mean_stab~log_mean_fun_balance+nutrients+temperature+richness+mean_int_rep_strength)
anova(M_int,lm_full)


M_int_syn<-lm(data=rep_aggr,mean_syn_Gross~log_mean_fun_balance+richness+nutrients+temperature+mean_int_rep_strength)


plot_richness_int<-ggplot(data=rep_aggr,aes(x=mean_int_rep_strength))+
  geom_histogram(bins=10)+
  facet_wrap(~nutrients)
plot_int_syn<-ggplot(data=rep_aggr,aes(y=mean_syn_Gross,x=mean_int_rep_strength))+
  geom_point()+
  geom_smooth(method="lm")+
  facet_wrap(~nutrients)

### sem

rep_aggr<-rep_aggr%>%mutate(richness=as.numeric(richness))




model_int <- '
  log_mean_stab ~ mean_syn_Gross
  +temperature
  +nutrients
  +log_mean_fun_balance
  
  
  mean_syn_Gross ~ log_mean_fun_balance
  +nutrients
  +richness
  +mean_int_rep_strength

  
'
fit_int <- sem(model_int, estimator="MLM",meanstructure = TRUE,data = rep_aggr)


summary(fit_int, standardized = TRUE, fit.measures = TRUE)



##### load interactions with T

```
## Check relationship of variables to mean biomass



```{r}
plot_biom_asyn<-ggplot(data=async_aggr,aes(x=synchrony_Gross,y=mean_biom))+
  geom_point()
  #geom_smooth(method="lm")

plot_biom_RD<-ggplot(data=async_aggr,aes(x=log10(avg_w_sumslopes),y=mean_biom))+
  geom_point()
  #geom_smooth(method="lm")

plot_sd_asyn<-ggplot(data=async_aggr,aes(x=synchrony_Gross,y=sd))+
  geom_point()
  #geom_smooth(method="lm")

plot_sd_RD<-ggplot(data=async_aggr,aes(x=log10(avg_w_sumslopes),y=sd))+
  geom_point()
  #geom_smooth(method="lm")


(plot_biom_asyn+plot_biom_RD)/
  (plot_sd_asyn+plot_sd_RD)

# create a temp stability group
async_aggr <- async_aggr %>% 
  mutate(CV_group = cut(1/CV, breaks = quantile(1/async_aggr$CV),include.lowest=T,labels=c("low","medium low","medium high","high")))


plot_syn_RD_CV<-ggplot(data=async_aggr,aes(x=log10(avg_w_sumslopes),y=synchrony_Gross,color=CV_group))+
  geom_point()

#-> Both show  similar relationship with mean biomass
```

##Poster

```{r}

### time series with nutrients==0.35 and T=22-25

poster_ts<-complete_ts%>%filter(nutrients=="0.35 g/L",
                                temperature=="22-25 °C",
                                sample_ID=="DLPS_Tmp_22_25_Nut0.35_3")
dark2_palette <- brewer.pal(n = 8, name = "Dark2")
selected_colors <- dark2_palette[2:5]

plot_poster_ts <- ggplot(data = poster_ts, aes(x = day, y = biomass, color = predict_spec)) +
  geom_line(linewidth=1.5) + ylab("Biomass")+xlab("Day")+
  theme_classic() +
  scale_color_manual(values = selected_colors) +
  theme(legend.position = "none",       
        axis.text.y = element_blank(),        # Remove the y-axis 
        axis.ticks.y = element_blank(),
        axis.title.x = element_text(size=16),
        axis.title.y=element_text(size=16))     

plot_poster_ts

cake_df<-poster_ts%>%group_by(predict_spec)%>%summarize(mean_biom=mean(biomass))


cake_df <- tibble::tibble(
  predict_spec = c("Dexiostoma", "Loxocephalus", "Paramecium", "Spirostomum"),
  mean_biom = c(0.000000843, 0.000000808, 0.00000520, 0.000000237)
)

# percentages
cake_df <- cake_df %>%
  mutate(percentage = mean_biom / sum(mean_biom) * 100)

# cake diagramm
plot_cake <- ggplot(cake_df, aes(x = "", y = percentage, fill = predict_spec)) +
  scale_fill_manual(values = selected_colors) +
  geom_bar(stat = "identity", width = 1) + 
  coord_polar(theta = "y") +                    # Titel und Legende
  theme_void()+
  labs(fill="Species")+
  theme(legend.direction = "vertical",legend.position="right",
        legend.title = element_text(size=16))# Verwende eine Farbpalette


plot_cake+plot_poster_ts+ plot_layout(ncol = 2,widths=c(1,2))   # Title for the combined plot
weight_plot <- wrap_elements(plot_cake + theme(plot.margin = margin( b = 150))) +  # Move the cake plot up
  plot_spacer() +
  plot_poster_ts + 
  plot_layout(ncol = 3, widths = c(1, 0.2, 1))

weight_plot

#ggsave("weight.pdf", plot = weight_plot, width = 10, height = 6)


# create temperature regime plot

regime1<-data.frame(temperature=rep(c(18,18,18,21,21,21),length.out=60),regime= "18-21 °C")
regime2<-data.frame(temperature=rep(c(22,22,22,25,25,25),length.out=60),regime= "22-25 °C")
regime3<-data.frame(temperature=rep(c(25,25,25,28,28,28),length.out=60),regime= "25-28 °C")
day=

df_Temp<-rbind(regime1,regime2,regime3)
df_Temp<-cbind(df_Temp,day=rep(1:60))

plot_temp<-ggplot(data=df_Temp,aes(x=day,y=temperature,linetype=regime))+
  geom_line()+  theme_classic() +
  ylab("Temperature °C")+xlab("Day")+
  labs(linetype="Temperature regime")+
  scale_y_continuous(breaks = c(28, 25, 22, 21, 18)) +
  theme(legend.direction = "vertical",legend.position="right",        
        axis.title.x = element_text(size=16),
        axis.title.y=element_text(size=16))


ggsave("temp_regime.pdf", plot = plot_temp, width = 8, height = 6)


############ Main results plot


balance_dd<-complete_aggr%>%select(abs_sum_slope,richness,CV)%>%
  cbind(type="no")%>%mutate(abs_sum_slope=abs_sum_slope/as.numeric(richness))
names(balance_dd)[names(balance_dd)=="abs_sum_slope"]<-"balance"

weighted_balance_dd<-complete_aggr%>%select(avg_w_sumslopes,richness,CV)%>%
  cbind(type="yes")
names(weighted_balance_dd)[names(weighted_balance_dd)=="avg_w_sumslopes"]<-"balance"

main_r_dd<-rbind(balance_dd,weighted_balance_dd)

plot_main <- main_r_dd %>%
  ggplot(aes(x = log10(balance), y = log10(1/CV), color = type)) +
  geom_point(aes(x = log10(balance), y = 1/CV)) +
  geom_smooth(method = "lm", aes(x = log10(balance), y = 1/CV)) +
  theme_classic() +
  scale_color_viridis_d(option = "inferno", begin = 0.3, end = 0.6) +
  scale_x_continuous(
    breaks = log10(c(0.0025,0.01, 0.04, 0.16, 0.64)),
    labels = c("0.0025","0.01", "0.04", "0.16", "0.64")
  ) +
  ylab(expression("temporal stability: "~log[10]("mean/s.d.")))+
  xlab(expression(log[10]("Balance"))) +
  theme(axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16),
        strip.background = element_rect(fill = "grey80", color = NA),  
    strip.text = element_text(size = 12)) +
  labs(color = "Weighted") +  
  facet_wrap(~richness, labeller = as_labeller(function(x) paste0("richness = ", x)))

ggsave("main_plot.pdf", plot = plot_main, width = 10, height = 6)



### linear model summary table

# Install and load the broom package if you haven't already

library(broom)

# Fit the model
lm_full<-lm(data=complete_aggr, log10(stability)~log10(abs_sum_slope)+richness+temperature+nutrients)

# Tidy up the summary
tidy_model <- tidy(lm_full_w)

# View the tidied model
tidy_model


library(broom)
library(writexl)



# Step 3: Tidy the model output using broom
model_summary <- tidy(lm_full)

# Step 4: Add significance stars for p-values
model_summary <- model_summary %>%
  mutate(Significance = case_when(
    p.value < 0.001 ~ "***",
    p.value < 0.01 ~ "**",
    p.value < 0.05 ~ "*",
    p.value < 0.1 ~ ".",
    TRUE ~ ""
  ))

# Step 5: Rename columns for a nicer presentation
names(model_summary) <- c("Term", "Estimate", "Std. Error", "t value", "p value", "Significance")

# Step 6: Write the summary to an Excel file
write_xlsx(model_summary, "full_model_summary.xlsx")


# Step 3: Tidy the model output using broom
model_summary <- tidy(lm_full_w)

# Step 4: Add significance stars for p-values
model_summary <- model_summary %>%
  mutate(Significance = case_when(
    p.value < 0.001 ~ "***",
    p.value < 0.01 ~ "**",
    p.value < 0.05 ~ "*",
    p.value < 0.1 ~ ".",
    TRUE ~ ""
  ))

# Step 5: Rename columns for a nicer presentation
names(model_summary) <- c("Term", "Estimate", "Std. Error", "t value", "p value", "Significance")

# Step 6: Write the summary to an Excel file
write_xlsx(model_summary, "full_model_w_summary.xlsx")



```

